### 2015-01-27 project1/checkpoint/main.tex
Getting latex build to work

\documentclass[a4paper,12pt]{article}

\begin{document}

HELLO WORLD \cite{pde_solution}
\bibliographystyle{plain}
\bibliography{main}

\end{document}

### 2015-01-27 project1/checkpoint/gulpfile.js

@@ -1,4 +1,30 @@
 var gulp = require('gulp');
+var async = require('async');
+var exec = require('child_process').exec;
+var del = require('del');
 
-gulp.task('default', function() {
+// Converts a string command into an async exec call, to be chained with async.series
+var asyncCommand = function(commandText) {
+    return function(callback) {
+        exec(commandText, callback);
+    };
+};
+
+gulp.task('clean', function(cb) {
+    del(['*.{aux,bbl,blg,dvi,log,pdf}'], cb);
+});
+
+
+gulp.task('default', ['clean'], function(cb) {
+    var latex = 'latex -halt-on-error main';
+    var bibtex = 'bibtex main';
+    var dvipdf = 'dvipdf main.dvi main.pdf';
+    
+    async.series([
+        asyncCommand(latex),
+        asyncCommand(bibtex),
+        asyncCommand(latex),
+        asyncCommand(latex),
+        asyncCommand(dvipdf)
+    ], cb);
 });

### 2015-01-28 project1/data_processing/gulpfile.js
Dumping the mean

@@ -60,9 +60,13 @@ gulp.task('northbound-input-distribution', function(taskDone) {
                 startOffsets.push( startTimes[i] - startTimes[i-1] );
             }
             
-            console.log(histogram(startOffsets, { bins: 25 }));
+            console.log(histogram(startOffsets, { bins: 20 }));
 
-
+            // calculate the mean
+            var meanInMillis = startOffsets.reduce(function(accumulator, item) {
+                return accumulator + item;
+            }, 0) / startOffsets.length;
+            console.log("Mean is " + meanInMillis/1000 + " seconds");
 
             taskDone();
         });

### 2015-01-28 project1/data_processing/gulpfile.js
Tweaking bin sizes

@@ -60,7 +60,7 @@ gulp.task('northbound-input-distribution', function(taskDone) {
                 startOffsets.push( startTimes[i] - startTimes[i-1] );
             }
             
-            console.log(histogram(startOffsets, { bins: 20 }));
+            console.log(histogram(startOffsets, { bins: 30 }));
 
             // calculate the mean
             var meanInMillis = startOffsets.reduce(function(accumulator, item) {

### 2015-01-28 project1/data_processing/histogram.js

@@ -18,12 +18,17 @@ module.exports = function(data, options) {
     var binWidth = (maxValue - minValue) / binCount;
 
     // initialize the bins to zero. From http://stackoverflow.com/a/13735425
-    var bins = Array.apply(null, new Array(binCount)).map(Number.prototype.valueOf, 0);
+    var bins = Array.apply(null, new Array(binCount + 1)).map(Number.prototype.valueOf, 0);
 
     // populate the bins
     data.forEach(function(item) {
         var binIndex = whichBin(item, minValue, binWidth);
         bins[binIndex] += 1;
+    });
+
+    // scale the bins to be probabilities
+    bins = bins.map(function(binValue) {
+        return binValue / N;
     });
 
     // now return the plotted result

### 2015-01-28 project1/data_processing/gulpfile.js
Writing my own histogram...

@@ -57,16 +57,16 @@ gulp.task('northbound-input-distribution', function(taskDone) {
             var N = startTimes.length;
             var startOffsets = [];
             for(var i = 1; i < N; i++) {
-                startOffsets.push( startTimes[i] - startTimes[i-1] );
+                startOffsets.push( (startTimes[i] - startTimes[i-1]) / 1000 );
             }
             
             console.log(histogram(startOffsets, { bins: 30 }));
 
             // calculate the mean
-            var meanInMillis = startOffsets.reduce(function(accumulator, item) {
+            var mean = startOffsets.reduce(function(accumulator, item) {
                 return accumulator + item;
             }, 0) / startOffsets.length;
-            console.log("Mean is " + meanInMillis/1000 + " seconds");
+            console.log("Mean is " + mean + " seconds");
 
             taskDone();
         });

### 2015-01-28 project1/data_processing/histogram.js

@@ -1,8 +1,24 @@
-var chart = require('ascii-chart');
+var histogram = require('ascii-histogram');
 
 var whichBin = function(value, binStart, binWidth) {
     return Math.floor((value - binStart) / binWidth);
 };
+
+var binName = function(binIndex, binStart, binWidth) {
+    var min = binIndex * binWidth + binStart;
+    var max = (binIndex + 1) * binWidth + binStart;
+    return min.toPrecision(4) + "-" + max.toPrecision(4);
+}
+
+var repeatString = function(string, percent, width) {
+    var count = Math.round(percent * width);
+    var outputString = "";
+    for(var i = 0; i < count; i++) {
+        outputString += string;
+    }
+
+    return outputString;
+}
 
 /**
  * Draws an ascii histogram by bucketing the values in the passed array
@@ -11,6 +27,7 @@ module.exports = function(data, options) {
     options = options || {};
 
     var binCount = options.bins || 10;
+    var width = options.width || 80;
     
     var N = data.length;
     var maxValue = Math.max.apply(Math, data);
@@ -26,11 +43,19 @@ module.exports = function(data, options) {
         bins[binIndex] += 1;
     });
 
-    // scale the bins to be probabilities
-    bins = bins.map(function(binValue) {
-        return binValue / N;
+    // Format as an object
+    var histogramFormattedBins = {};
+    bins.forEach(function(count, index) {
+        histogramFormattedBins[ binName(index, minValue, binWidth) ] = count / N;
     });
 
-    // now return the plotted result
-    return chart(bins);
+    var outputString = "";
+    for(var label in histogramFormattedBins) {
+        outputString += label + ' | ';
+        outputString += repeatString('=', histogramFormattedBins[label], width);
+        outputString += "\n";
+    }
+
+    return outputString;
+
 };

### 2015-01-29 project1/data_processing/gulpfile.js
Playing with data processing code

@@ -52,21 +52,31 @@ gulp.task('northbound-input-distribution', function(taskDone) {
         }))
         .on('finish', function() {
             // sort the start times and calculate the differences to map to a Exponential
-            startTimes = startTimes.map(Number)
-            startTimes.sort();
+            startTimes.sort(function(a, b) {
+                return Number(a) - Number(b);
+            });
             var N = startTimes.length;
             var startOffsets = [];
             for(var i = 1; i < N; i++) {
                 startOffsets.push( (startTimes[i] - startTimes[i-1]) / 1000 );
             }
             
+            console.log(startOffsets);
             console.log(histogram(startOffsets, { bins: 30 }));
 
             // calculate the mean
             var mean = startOffsets.reduce(function(accumulator, item) {
                 return accumulator + item;
             }, 0) / startOffsets.length;
-            console.log("Mean is " + mean + " seconds");
+            console.log("Raw Mean is " + mean + " seconds");
+
+            var reducedOffsets = startOffsets.filter(function(value) {
+                return value < 40;
+            });
+            var reducedMean = reducedOffsets.reduce(function(accumulator, item) {
+                return accumulator + item;
+            }, 0) / reducedOffsets.length;
+            console.log('Reduced mean is ' + reducedMean);
 
             taskDone();
         });

### 2015-01-29 project1/data_processing/histogram.js

@@ -43,6 +43,8 @@ module.exports = function(data, options) {
         bins[binIndex] += 1;
     });
 
+    console.log(bins);
+
     // Format as an object
     var histogramFormattedBins = {};
     bins.forEach(function(count, index) {

### 2015-01-29 project1/data_processing/exponential.js
Adding a helper module for the exponential distribution

@@ -1,0 +1,9 @@
+
+module.exports = {
+    pdf: function(x, lambda) {
+        return lambda * Math.exp(-lambda * x);
+    },
+    cdf: function(x, lambda) {
+        return 1 - Math.exp(-lambda * x);
+    }
+};

### 2015-01-29 project1/data_processing/gulpfile.js
Adding a mean function to clean up a little

@@ -1,6 +1,7 @@
 var gulp = require('gulp');
 var csv = require('csv');
 var fs = require('fs');
+var mean = require('./mean.js');
 var histogram = require('./histogram.js');
 
 /**
@@ -65,17 +66,13 @@ gulp.task('northbound-input-distribution', function(taskDone) {
             console.log(histogram(startOffsets, { bins: 30 }));
 
             // calculate the mean
-            var mean = startOffsets.reduce(function(accumulator, item) {
-                return accumulator + item;
-            }, 0) / startOffsets.length;
-            console.log("Raw Mean is " + mean + " seconds");
+            var meanOffset = mean(startOffsets);
+            console.log("Raw Mean is " + meanOffset + " seconds");
 
             var reducedOffsets = startOffsets.filter(function(value) {
                 return value < 40;
             });
-            var reducedMean = reducedOffsets.reduce(function(accumulator, item) {
-                return accumulator + item;
-            }, 0) / reducedOffsets.length;
+            var reducedMean = mean(reducedOffsets);
             console.log('Reduced mean is ' + reducedMean);
 
             taskDone();

### 2015-01-29 project1/data_processing/mean.js

@@ -1,0 +1,5 @@
+module.exports = function(items) {
+    return items.reduce(function(accumulator, item) {
+        return accumulator + item;
+    }, 0) / items.length;
+};

### 2015-01-29 project1/data_processing/array_interval.js
A little more cleanup

@@ -1,0 +1,15 @@
+module.exports = function(items) {
+    // first sort the items numerically
+    items.sort(function(a, b) {
+        return Number(a) - Number(b);
+    });
+
+    // now build the offsets list
+    var N = items.length;
+    var offsets = [];
+    for(var i = 1; i < N; i++) {
+        offsets.push( items[i] - items[i-1] );
+    }
+
+    return offsets;
+};

### 2015-01-29 project1/data_processing/gulpfile.js

@@ -1,6 +1,7 @@
 var gulp = require('gulp');
 var csv = require('csv');
 var fs = require('fs');
+var array_interval = require('./array_interval.js');
 var mean = require('./mean.js');
 var histogram = require('./histogram.js');
 
@@ -52,15 +53,8 @@ gulp.task('northbound-input-distribution', function(taskDone) {
             return row;
         }))
         .on('finish', function() {
-            // sort the start times and calculate the differences to map to a Exponential
-            startTimes.sort(function(a, b) {
-                return Number(a) - Number(b);
-            });
-            var N = startTimes.length;
-            var startOffsets = [];
-            for(var i = 1; i < N; i++) {
-                startOffsets.push( (startTimes[i] - startTimes[i-1]) / 1000 );
-            }
+            // calculate the intervals between arrivals
+            var startOffsets = array_interval(startTimes);
             
             console.log(startOffsets);
             console.log(histogram(startOffsets, { bins: 30 }));

### 2015-01-29 project1/data_processing/gulpfile.js
Making the default task more useful

@@ -73,10 +73,4 @@ gulp.task('northbound-input-distribution', function(taskDone) {
         });
 });
 
-gulp.task('input-distribution', function() {
-    return keepEntrances()
-        .pipe(csv.stringify())
-        .pipe(process.stdout);
-});
-
-gulp.task('default', ['input-distribution']);
+gulp.task('default', ['northbound-input-distribution']);

### 2015-01-29 project1/data_processing/bin.js
Moving binning to a new file

@@ -1,0 +1,32 @@
+var Bin = function(items, binCount) {
+    this.rawItems = items;
+    this.binCount = binCount || 25;
+
+    // Initialize bins
+    this.bins = [];
+    for(var i = 0; i < this.binCount; i++) {
+        this.bins.push([]);
+    }
+
+    var N = items.length;
+    this.maxValue = Math.max.apply(Math, items);
+    this.minValue = Math.min.apply(Math, items);
+    this.binWidth = (this.maxValue - this.minValue ) / this.binCount;
+    var that = this;
+    items.forEach(function(item) {
+        var binIndex = that.whichBin(item, that.minValue, that.binWidth);
+        that.bins[binIndex].push(item);
+    });
+};
+
+Bin.prototype.whichBin = function(value) {
+    return Math.floor((value - this.minValue) / this.binWidth);
+};
+
+Bin.prototype.binName = function(index) {
+    var min = index * this.binWidth + this.minValue;
+    var max = min + this.binWidth;
+    return min.toPrecision(4) + "-" + max.toPrecision(4);
+};
+
+module.exports = Bin;

### 2015-01-29 project1/data_processing/histogram.js

@@ -1,14 +1,5 @@
 var histogram = require('ascii-histogram');
-
-var whichBin = function(value, binStart, binWidth) {
-    return Math.floor((value - binStart) / binWidth);
-};
-
-var binName = function(binIndex, binStart, binWidth) {
-    var min = binIndex * binWidth + binStart;
-    var max = (binIndex + 1) * binWidth + binStart;
-    return min.toPrecision(4) + "-" + max.toPrecision(4);
-}
+var Bin = require('./bin.js');
 
 var repeatString = function(string, percent, width) {
     var count = Math.round(percent * width);
@@ -26,30 +17,15 @@ var repeatString = function(string, percent, width) {
 module.exports = function(data, options) {
     options = options || {};
 
-    var binCount = options.bins || 10;
     var width = options.width || 80;
     
-    var N = data.length;
-    var maxValue = Math.max.apply(Math, data);
-    var minValue = Math.min.apply(Math, data);
-    var binWidth = (maxValue - minValue) / binCount;
-
-    // initialize the bins to zero. From http://stackoverflow.com/a/13735425
-    var bins = Array.apply(null, new Array(binCount + 1)).map(Number.prototype.valueOf, 0);
-
-    // populate the bins
-    data.forEach(function(item) {
-        var binIndex = whichBin(item, minValue, binWidth);
-        bins[binIndex] += 1;
-    });
-
-    console.log(bins);
+    var bin = new Bin(data, options.bins);
 
     // Format as an object
     var histogramFormattedBins = {};
-    bins.forEach(function(count, index) {
-        histogramFormattedBins[ binName(index, minValue, binWidth) ] = count / N;
-    });
+    for(var i = 0; i < bin.binCount; i++) {
+        histogramFormattedBins[ bin.binName(i) ] = bin.bins[i].length / bin.rawItems.length;
+    }
 
     var outputString = "";
     for(var label in histogramFormattedBins) {
@@ -59,5 +35,4 @@ module.exports = function(data, options) {
     }
 
     return outputString;
-
 };

### 2015-01-29 project1/data_processing/expectedExponentialFrequencies.js
Wiring up everything. Bins is still broken

@@ -1,0 +1,17 @@
+var exponential = require('./exponential.js');
+
+var probabilityBetween = function(a, b, lambda) {
+    return exponential.cdf(b, lambda) - exponential.cdf(a, lambda);
+};
+
+module.exports = function(binWidth, binCount, lambda) {
+    var frequencies  = [];
+
+    for(var i = 0; i < binCount; i++) {
+        var leftBoundary = i * binWidth;
+        var rightBoundary = (i + 1) * binWidth;
+        frequencies.push(probabilityBetween(leftBoundary, rightBoundary, lambda));
+    }
+
+    return frequencies;
+};

### 2015-01-29 project1/data_processing/gulpfile.js

@@ -3,6 +3,7 @@ var csv = require('csv');
 var fs = require('fs');
 var array_interval = require('./array_interval.js');
 var mean = require('./mean.js');
+var performGoodnessOfFit = require('./performGoodnessOfFit.js');
 var histogram = require('./histogram.js');
 
 /**
@@ -53,21 +54,12 @@ gulp.task('northbound-input-distribution', function(taskDone) {
             return row;
         }))
         .on('finish', function() {
-            // calculate the intervals between arrivals
-            var startOffsets = array_interval(startTimes);
-            
-            console.log(startOffsets);
-            console.log(histogram(startOffsets, { bins: 30 }));
-
-            // calculate the mean
-            var meanOffset = mean(startOffsets);
-            console.log("Raw Mean is " + meanOffset + " seconds");
-
-            var reducedOffsets = startOffsets.filter(function(value) {
-                return value < 40;
+            // calculate the intervals between arrivals (in seconds)
+            var startOffsets = array_interval(startTimes).map(function(interval) {
+                return interval / 1000;
             });
-            var reducedMean = mean(reducedOffsets);
-            console.log('Reduced mean is ' + reducedMean);
+            
+            console.log("Chi-Squared Test Probability: " + performGoodnessOfFit(startOffsets, 25));
 
             taskDone();
         });

### 2015-01-29 project1/data_processing/performGoodnessOfFit.js

@@ -1,0 +1,22 @@
+var expectedExponentialFrequencies = require('./expectedExponentialFrequencies.js');
+var mean = require('./mean.js');
+var Bin = require('./bin.js');
+var chiSquaredTest = require('chi-squared-test');
+
+module.exports = function(rawData, binCount) {
+    var binnedData = new Bin(rawData, binCount);
+
+    // The maximum likelihood lambda value 1 / mean
+    var mleLambda = 1 / mean(rawData);
+
+    // the expected frequencies
+    var expectations = expectedExponentialFrequencies(binnedData.binWidth, binCount, mleLambda);
+
+    // the actual frequencies
+    var N = rawData.length;
+    var actualFrequencies = binnedData.bins.map(function(bin) {
+        return bin.length / N;
+    });
+
+    return chiSquaredTest(actualFrequencies, expectations, 2);
+};

### 2015-01-29 project1/data_processing/bin.js
Getting bin to work

@@ -9,7 +9,7 @@ var Bin = function(items, binCount) {
     }
 
     var N = items.length;
-    this.maxValue = Math.max.apply(Math, items);
+    this.maxValue = Math.max.apply(Math, items) * 1.05;
     this.minValue = Math.min.apply(Math, items);
     this.binWidth = (this.maxValue - this.minValue ) / this.binCount;
     var that = this;

### 2015-01-29 project1/data_processing/performGoodnessOfFit.js
Fixing goodness of fit to use frequencies instead

@@ -9,13 +9,16 @@ module.exports = function(rawData, binCount) {
     // The maximum likelihood lambda value 1 / mean
     var mleLambda = 1 / mean(rawData);
 
+    var N = rawData.length;
+
     // the expected frequencies
-    var expectations = expectedExponentialFrequencies(binnedData.binWidth, binCount, mleLambda);
+    var expectations = expectedExponentialFrequencies(binnedData.binWidth, binCount, mleLambda).map(function(percent) {
+        return percent * N;
+    });
 
     // the actual frequencies
-    var N = rawData.length;
     var actualFrequencies = binnedData.bins.map(function(bin) {
-        return bin.length / N;
+        return bin.length;
     });
 
     return chiSquaredTest(actualFrequencies, expectations, 2);

### 2015-01-29 project1/data_processing/gulpfile.js
Using new format for chi-squared, trying sublist?

@@ -59,7 +59,13 @@ gulp.task('northbound-input-distribution', function(taskDone) {
                 return interval / 1000;
             });
             
-            console.log("Chi-Squared Test Probability: " + performGoodnessOfFit(startOffsets, 25));
+            var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
+            console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
+
+            // try skipping the last quarter of points?
+            var subListLength = Math.round(startOffsets.length * 0.25);
+            chiSquaredResults = performGoodnessOfFit(startOffsets.slice(0, subListLength), 25);
+            console.log("Skipped points Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
 
             taskDone();
         });

### 2015-01-29 project1/data_processing/gulpfile.js
Removing some skipping. I think I'm looking at the wrong points?

@@ -58,14 +58,11 @@ gulp.task('northbound-input-distribution', function(taskDone) {
             var startOffsets = array_interval(startTimes).map(function(interval) {
                 return interval / 1000;
             });
-            
-            var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
-            console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
 
-            // try skipping the last quarter of points?
-            var subListLength = Math.round(startOffsets.length * 0.25);
-            chiSquaredResults = performGoodnessOfFit(startOffsets.slice(0, subListLength), 25);
-            console.log("Skipped points Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
+            var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
+            console.log('N = ' + startOffsets.length);
+            console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
+            console.log(histogram(startOffsets, { }));
 
             taskDone();
         });

### 2015-01-29 project1/data_processing/gulpfile.js
Trying out cars in the intersection

@@ -41,8 +41,8 @@ gulp.task('northbound-input-distribution', function(taskDone) {
 
     getCsv()
         .pipe(csv.transform(function(row) {
-            // Filter only rows that are northbound, and in the correct section of road
-            if(row[17] != '3' || row[18] != '2') {
+            // Filter only rows that are northbound, and in the intersection
+            if(row[16] != '3' || row[18] != '2') {
                 return;
             }
 

### 2015-01-31 project1/checkpoint/main.tex
Stubbing out basic structure for paper
Adding the first draft of the problem description

@@
 
-HELLO WORLD \cite{pde_solution}
+\title{Project 1 Checkpoint}
+\author{Chip Bell}
+\date{February 6th, 2015}
+\maketitle
+
+\section{Problem Description}
+Traffic simulation is becoming increasing important due to the expansion and growth of cities. Long-standing
+infrastructure is forced to adapt in order to accomodate greater amounts of traffic, despite the fact that in many
+cases the roads cannot be modified. This can be a result of legislation, existing buildings, and private property among
+other things. So as a result, cities must instead optimize the variables they can control. In this case, we consider
+signal light timing for the 12th St. and Peachtree Street intersection in only the north and south directions due to
+time constraints, and the sheer difficulty with modeling traffic in general. In particular, we explore the effects that
+light timing has on traffic flow, given a source input distribution of cars.
+
+\section{The Conceptual Model}
+Conceptual Model
+
+\section{Simulation Software Architecture}
+Simulation Software
+
+\section{Current Build}
+Current Build
+
+\cite{pde_solution}
 \bibliographystyle{plain}

### 2015-01-31 project1/checkpoint/main.tex
Adding bibtex entry for main report
Stubbing out some subsections and adding some assumptions

@@
 \section{The Conceptual Model}
-Conceptual Model
+In order to create a representative software simulation of the aforementioned system, we first begin by designing a
+conceptual version of this model from which we can plan our implementation. Let's consider a couple of key aspects
+categorically:
+
+\subsection{Simulation Input}
+
+\subsection{Simulation Output}
+
+\subsection{Content}
+
+\subsection{Simplifications and Assumptions}
+Probably the most obvious "flaw" in such a design are the simplifications made. Details can certainly be left out
+purposely for many reasons such as time constraints, lack of supporting data to model a extra feature, or simply
+unneccesary complexity.
+
+In this case, certainly the most prominent would be the ``isolation'' of the model. We are only considering an
+idealized random input distribution of cars from both North and South of the 12th St. intersection. However, this is
+certainly not the case, since this input distribution is affected by countless other factors like the surrounding
+lights, cars passing along 12th at that intersection, and even simply time of day. We assume in this simulation is
+perpetually between 4:00PM to 4:15PM in Atlanta (a frightening thought indeed), simply due to the fact that we only
+have data for that time range.
 
@@
 
-\cite{pde_solution}
+\cite{cts12}
 \bibliographystyle{plain}

### 2015-01-31 project1/checkpoint/main.tex
Adding a few more details
Saving WIP

@@
 \documentclass[a4paper,12pt]{article}
+
+\usepackage{url, hyperref}
 
@@
 other things. So as a result, cities must instead optimize the variables they can control. In this case, we consider
-signal light timing for the 12th St. and Peachtree Street intersection in only the north and south directions due to
-time constraints, and the sheer difficulty with modeling traffic in general. In particular, we explore the effects that
-light timing has on traffic flow, given a source input distribution of cars.
+signal light timing for the 12th St. and Peachtree Street intersection due to
+time constraints, and the sheer difficulty with modeling traffic in general between multiple intersections. In particular,
+we explore the effects that light timing has on traffic flow, given a source input distribution of cars.
 
@@
 In order to create a representative software simulation of the aforementioned system, we first begin by designing a
-conceptual version of this model from which we can plan our implementation. Let's consider a couple of key aspects
-categorically:
+conceptual version of this model from which we can plan our implementation. Data from \cite{cts12} was used extensively % TODO: Put this somewhere else
+for modeling purposes in this project. Our first step will be to identify the entities present in our model, and then consider their interactions (or the
+inherent ``activies'') within the simulation.
+
+The first entity, unsurprisingly, is a single car. Our representation of a car is clearly less detailed than the real
+thing, but rather focuses on variables that influence it's travel time from its starting location to destination, along
+with the variables that influence neighboring cars. For our simulation, we'll track the position and destination of the
+cars, along with the size of the car since that can influence the position of cars behind it. Also, we will track the
+current velocity of the car
+
+Another entity to consider is the intersection itself. It's changing in time constantly, and those changes affect
+every entity within the simulation. The most important parameters to consider here are the light timings. This
+intersection's timings are time-based, rather than sensor-based, which can be seen in the data set provided with the
+project description and is also available online \cite{ngsim}. We assume that the current light timing provided here
+are optimal, but because we'll have a computer simulation in the end we have free reign in experimenting with these
+values and observing the outcome. Perhaps the timings are \emph{not} optimal. Perhaps, our model grossly oversimplifies
+some crucial interaction within the system. This is perhaps the most exciting part of the project.
+
+From the intersection entity, we can glean an important interaction that occurs: The changing of light signal. When a
+signal changes, each car in-queue will be forced to change it's state, whether it's waiting or currently in motion.
 
 \subsection{Simulation Input}
+The two main sources of input into this simulation are the cars entering the section of Peachtree St with which we are
+concerned, and the light timing of the system. The combination of these two variables are what influence the smoothness
+of traffic flow to the actual drivers in the system.
+
+We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
+consider the car arrival time, and car length when modeling cars. Due to the discrete nature of the application, we'll
+ignore some of the more "continuous" aspects, such car velocity and acceleration and encompass them into calculation
+of event times.
 
 \subsection{Simulation Output}
-
-\subsection{Content}
 
@@
 perpetually between 4:00PM to 4:15PM in Atlanta (a frightening thought indeed), simply due to the fact that we only
-have data for that time range.
+have data for that time range. All in all, we assume the conditions were fairly identical to that in which the 
+aforementioned CTS study was made.
 

### 2015-02-01 project1/data_processing/csvFilter.js
Cleaning up code a little

@@ -1,0 +1,30 @@
+var csv = require('csv');
+var fs = require('fs');
+
+/**
+ * Helper method for querying the CSV file
+ */
+var getCsv = function() {
+    // return a stream filtered and transformed into a new format
+    return fs.createReadStream('data.csv')
+        .pipe(csv.parse());
+};
+
+/**
+ * Provides an easy way to hook into filtering the CSV file
+ */
+module.exports = function(filters, finishCode) {
+
+    var filterCount = filters.length;
+
+    var stream = getCsv();
+
+    for(var i = 0; i < filterCount; i++) {
+        stream = stream.pipe(csv.transform(filters[i]));
+    }
+
+    // now add the end stream
+    stream.on('finish', finishCode);
+
+    return stream;
+};

### 2015-02-01 project1/data_processing/gulpfile.js

@@ -1,19 +1,10 @@
 var gulp = require('gulp');
-var csv = require('csv');
-var fs = require('fs');
+var csvFilter = require('./csvFilter.js');
+var ngsimFilters = require('./ngsimFilters.js');
 var array_interval = require('./array_interval.js');
 var mean = require('./mean.js');
 var performGoodnessOfFit = require('./performGoodnessOfFit.js');
 var histogram = require('./histogram.js');
-
-/**
- * Helper method for querying the CSV file
- */
-var getCsv = function() {
-    // return a stream filtered and transformed into a new format
-    return fs.createReadStream('data.csv')
-        .pipe(csv.parse());
-};
 
 /**
  * Helper method for only getting the first appearances of the cars
@@ -32,40 +23,45 @@ var keepFirst = function() {
         return row;
     };
 
-    return csv.transform(firstTracker);
+    return firstTracker;
 };
 
 gulp.task('northbound-input-distribution', function(taskDone) {
 
     var startTimes = [];
 
-    getCsv()
-        .pipe(csv.transform(function(row) {
-            // Filter only rows that are northbound, and in the intersection
-            if(row[16] != '3' || row[18] != '2') {
-                return;
-            }
+    var filters = [];
 
-            return row;
-        }))
-        .pipe(keepFirst())
-        .pipe(csv.transform(function(row) {
-            startTimes.push(Number(row[3]));
-            return row;
-        }))
-        .on('finish', function() {
-            // calculate the intervals between arrivals (in seconds)
-            var startOffsets = array_interval(startTimes).map(function(interval) {
-                return interval / 1000;
-            });
+    // Only northbound
+    filters.push(ngsimFilters.direction(ngsimFilters.direction.NORTH));
 
-            var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
-            console.log('N = ' + startOffsets.length);
-            console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
-            console.log(histogram(startOffsets, { }));
+    // only through our intersection
+    filters.push(ngsimFilters.intersection('3'));
 
-            taskDone();
+    // only keep the first occurrences
+    filters.push(keepFirst());
+
+    // track the start times
+    filters.push(function(row) {
+        startTimes.push(Number(row[3]));
+        return row;
+    });
+
+    var finish = function() {
[…]

### 2015-02-01 project1/data_processing/ngsimFilters.js

@@ -1,0 +1,39 @@
+/**
+ * Basic function for building a filter for a given row on a certain column
+ */
+var filterFactory = function(columnIndex, value) {
+    return function(row) {
+        if(row[columnIndex] != value) {
+            return;
+        }
+
+        return row;
+    };
+};
+
+/**
+ * Factory for building a filter for direction
+ */
+var filterDirection = function(direction) {
+    return filterFactory(18, direction); // 18 is direction
+};
+
+filterDirection.EAST = '1';
+filterDirection.NORTH = '2';
+filterDirection.WEST = '3';
+filterDirection.SOUTH = '4';
+
+/**
+ * Filters by intersection
+ */
+var filterIntersection = function(intersection) {
+    return filterFactory(16, intersection);
+};
+
+// TODO: Name the intersections, so users have something to work with
+
+module.exports = {
+    direction: filterDirection,
+    intersection: filterIntersection,
+    factory: filterFactory
+};

### 2015-02-01 project1/data_processing/array_interval.js
Adding in some comments

@@ -1,3 +1,7 @@
+/**
+ * Helper function to take an array and take the difference with
+ * previous elements. For instance [1, 2, 4, 4] would become [1, 2, 0]
+ */
 module.exports = function(items) {
     // first sort the items numerically
     items.sort(function(a, b) {

### 2015-02-01 project1/data_processing/bin.js

@@ -1,3 +1,6 @@
+/**
+ * Helper class for binning a list of items (a frequency distribution calculator)
+ */
 var Bin = function(items, binCount) {
     this.rawItems = items;
     this.binCount = binCount || 25;
@@ -19,10 +22,16 @@ var Bin = function(items, binCount) {
     });
 };
 
+/**
+ * Returns the bin index that a value would fall into
+ */
 Bin.prototype.whichBin = function(value) {
     return Math.floor((value - this.minValue) / this.binWidth);
 };
 
+/**
+ * Determines a "pretty name" for a bin
+ */
 Bin.prototype.binName = function(index) {
     var min = index * this.binWidth + this.minValue;
     var max = min + this.binWidth;

### 2015-02-01 project1/data_processing/csvFilter.js

@@ -19,11 +19,12 @@ module.exports = function(filters, finishCode) {
 
     var stream = getCsv();
 
+    // tag on all of the filters
     for(var i = 0; i < filterCount; i++) {
         stream = stream.pipe(csv.transform(filters[i]));
     }
 
-    // now add the end stream
+    // now add the end stream listener
     stream.on('finish', finishCode);
 
     return stream;

### 2015-02-01 project1/data_processing/expectedExponentialFrequencies.js

@@ -1,9 +1,19 @@
 var exponential = require('./exponential.js');
 
+/**
+ * calculates the integral of the exponential pdf between two values
+ */
 var probabilityBetween = function(a, b, lambda) {
     return exponential.cdf(b, lambda) - exponential.cdf(a, lambda);
 };
 
+/**
+ * Calculates the expected probabilities for a set of bins provided by:
+ *
+ * @param binWidth The width of each bin
+ * @param binCount The number of bins
+ * @param lambda   The lambda to use
+ */
 module.exports = function(binWidth, binCount, lambda) {
     var frequencies  = [];
 

### 2015-02-01 project1/data_processing/exponential.js

@@ -1,4 +1,6 @@
-
+/**
+ * The exponential distribution
+ */
 module.exports = {
     pdf: function(x, lambda) {
         return lambda * Math.exp(-lambda * x);

### 2015-02-01 project1/data_processing/histogram.js

@@ -1,6 +1,13 @@
 var histogram = require('ascii-histogram');
 var Bin = require('./bin.js');
 
+/**
+ * Repeats a string
+ *
+ * @param string  The string to repeat
+ * @param percent The percentage of the full width to stretch
+ * @param width   The size that 100% means
+ */
 var repeatString = function(string, percent, width) {
     var count = Math.round(percent * width);
     var outputString = "";

### 2015-02-01 project1/data_processing/mean.js

@@ -1,3 +1,6 @@
+/**
+ * Calculates the average of a list.
+ */
 module.exports = function(items) {
     return items.reduce(function(accumulator, item) {
         return accumulator + item;

### 2015-02-01 project1/data_processing/performGoodnessOfFit.js

@@ -3,6 +3,10 @@ var mean = require('./mean.js');
 var Bin = require('./bin.js');
 var chiSquaredTest = require('chi-squared-test');
 
+/**
+ * Performs an exponential goodness-of-fit test to determine how good a fit some raw data is. Essentially performs a
+ * chi-squared test
+ */
 module.exports = function(rawData, binCount) {
     var binnedData = new Bin(rawData, binCount);
 

### 2015-02-01 project1/data_processing/gulpfile.js
Pulling first keeper to a better home

@@ -6,26 +6,6 @@ var mean = require('./mean.js');
 var performGoodnessOfFit = require('./performGoodnessOfFit.js');
 var histogram = require('./histogram.js');
 
-/**
- * Helper method for only getting the first appearances of the cars
- */
-var keepFirst = function() {
-    var seenCars = {};
-    var firstTracker = function(row) {
-        var car_id = row[0];
-
-        // if we've seen the car skip it
-        if(seenCars[car_id]) {
-            return null;
-        }
-
-        seenCars[car_id] = true;
-        return row;
-    };
-
-    return firstTracker;
-};
-
 gulp.task('northbound-input-distribution', function(taskDone) {
 
     var startTimes = [];
@@ -39,7 +19,7 @@ gulp.task('northbound-input-distribution', function(taskDone) {
     filters.push(ngsimFilters.intersection('3'));
 
     // only keep the first occurrences
-    filters.push(keepFirst());
+    filters.push(ngsimFilters.firstKeeper());
 
     // track the start times
     filters.push(function(row) {

### 2015-02-01 project1/data_processing/ngsimFilters.js

@@ -31,9 +31,31 @@ var filterIntersection = function(intersection) {
 };
 
 // TODO: Name the intersections, so users have something to work with
+//
+
+/**
+ * Factory for a filter that only keeps the first occurrence of car
+ */
+var keepFirst = function() {
+    var seenCars = {};
+    var firstTracker = function(row) {
+        var car_id = row[0];
+
+        // if we've seen the car skip it
+        if(seenCars[car_id]) {
+            return null;
+        }
+
+        seenCars[car_id] = true;
+        return row;
+    };
+
+    return firstTracker;
+};
 
 module.exports = {
     direction: filterDirection,
     intersection: filterIntersection,
+    firstKeeper: keepFirst,
     factory: filterFactory
 };

### 2015-02-01 project1/data_processing/gulpfile.js
Making task generic

@@ -1,47 +1,10 @@
 var gulp = require('gulp');
-var csvFilter = require('./csvFilter.js');
 var ngsimFilters = require('./ngsimFilters.js');
-var array_interval = require('./array_interval.js');
-var mean = require('./mean.js');
-var performGoodnessOfFit = require('./performGoodnessOfFit.js');
-var histogram = require('./histogram.js');
+var inputDistributionTaskFactory = require('./inputDistributionTaskFactory.js');
 
-gulp.task('northbound-input-distribution', function(taskDone) {
-
-    var startTimes = [];
-
-    var filters = [];
-
-    // Only northbound
-    filters.push(ngsimFilters.direction(ngsimFilters.direction.NORTH));
-
-    // only through our intersection
-    filters.push(ngsimFilters.intersection('3'));
-
-    // only keep the first occurrences
-    filters.push(ngsimFilters.firstKeeper());
-
-    // track the start times
-    filters.push(function(row) {
-        startTimes.push(Number(row[3]));
-        return row;
-    });
-
-    var finish = function() {
-        // calculate the intervals between arrivals (in seconds)
-        var startOffsets = array_interval(startTimes).map(function(interval) {
-            return interval / 1000;
-        });
-
-        var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
-        console.log('N = ' + startOffsets.length);
-        console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
-        console.log(histogram(startOffsets, { }));
-
-        taskDone();
-    };
-
-    csvFilter(filters, finish);
-});
+gulp.task(
+    'northbound-input-distribution',
+    inputDistributionTaskFactory(ngsimFilters.direction.NORTH, '3')
+);
 
 gulp.task('default', ['northbound-input-distribution']);

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js

@@ -1,0 +1,48 @@
+var csvFilter = require('./csvFilter.js');
+var ngsimFilters = require('./ngsimFilters.js');
+var array_interval = require('./array_interval.js');
+var mean = require('./mean.js');
+var performGoodnessOfFit = require('./performGoodnessOfFit.js');
+var histogram = require('./histogram.js');
+
+/**
+ * Factory function for building a gulp task for filtering the NGSIM data for a certain direction and section of the road
+ */
+module.exports = function(direction, section) {
+    return function(taskDone) {
+        var startTimes = [];
+
+        var filters = [];
+
+        // Only northbound
+        filters.push(ngsimFilters.direction(direction));
+
+        // only through our intersection
+        filters.push(ngsimFilters.intersection(section));
+
+        // only keep the first occurrences
+        filters.push(ngsimFilters.firstKeeper());
+
+        // track the start times
+        filters.push(function(row) {
+            startTimes.push(Number(row[3]));
+            return row;
+        });
+
+        var finish = function() {
+            // calculate the intervals between arrivals (in seconds)
+            var startOffsets = array_interval(startTimes).map(function(interval) {
+                return interval / 1000;
+            });
+
+            var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
+            console.log('N = ' + startOffsets.length);
+            console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
+            console.log(histogram(startOffsets, { }));
+
+            taskDone();
+        };
+
+        csvFilter(filters, finish);
+    };
+};

### 2015-02-01 project1/data_processing/gulpfile.js
Shortening names up

@@ -1,10 +1,7 @@
 var gulp = require('gulp');
-var ngsimFilters = require('./ngsimFilters.js');
-var inputDistributionTaskFactory = require('./inputDistributionTaskFactory.js');
+var ngsim = require('./ngsimFilters.js');
+var inputDistribution = require('./inputDistributionTaskFactory.js');
 
-gulp.task(
-    'northbound-input-distribution',
-    inputDistributionTaskFactory(ngsimFilters.direction.NORTH, '3')
-);
+gulp.task('northbound', inputDistribution(ngsim.direction.NORTH, '3'));
 
-gulp.task('default', ['northbound-input-distribution']);
+gulp.task('default', ['northbound']);

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js
Fixing some wording

@@ -6,9 +6,9 @@ var performGoodnessOfFit = require('./performGoodnessOfFit.js');
 var histogram = require('./histogram.js');
 
 /**
- * Factory function for building a gulp task for filtering the NGSIM data for a certain direction and section of the road
+ * Factory function for building a gulp task for filtering the NGSIM data for a certain direction and intersection of the road
  */
-module.exports = function(direction, section) {
+module.exports = function(direction, intersection) {
     return function(taskDone) {
         var startTimes = [];
 

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js
Forgot one

@@ -18,7 +18,7 @@ module.exports = function(direction, intersection) {
         filters.push(ngsimFilters.direction(direction));
 
         // only through our intersection
-        filters.push(ngsimFilters.intersection(section));
+        filters.push(ngsimFilters.intersection(intersection));
 
         // only keep the first occurrences
         filters.push(ngsimFilters.firstKeeper());

### 2015-02-01 project1/data_processing/gulpfile.js
Using an options object rather than parameters

@@ -2,6 +2,14 @@ var gulp = require('gulp');
 var ngsim = require('./ngsimFilters.js');
 var inputDistribution = require('./inputDistributionTaskFactory.js');
 
-gulp.task('northbound', inputDistribution(ngsim.direction.NORTH, '3'));
+gulp.task('northbound', inputDistribution({
+    direction: ngsim.direction.NORTH,
+    intersection: '3'
+}));
+
+gulp.task('southbound', inputDistribution({
+    direction: ngsim.direction.SOUTH,
+    intersection: '3'
+}));
 
 gulp.task('default', ['northbound']);

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js

@@ -8,17 +8,21 @@ var histogram = require('./histogram.js');
 /**
  * Factory function for building a gulp task for filtering the NGSIM data for a certain direction and intersection of the road
  */
-module.exports = function(direction, intersection) {
+module.exports = function(options) {
     return function(taskDone) {
         var startTimes = [];
 
         var filters = [];
 
-        // Only northbound
-        filters.push(ngsimFilters.direction(direction));
+        // Only one direction
+        if(options.direction) {
+            filters.push(ngsimFilters.direction(options.direction));
+        }
 
         // only through our intersection
-        filters.push(ngsimFilters.intersection(intersection));
+        if(options.intersection) {
+            filters.push(ngsimFilters.intersection(options.intersection));
+        }
 
         // only keep the first occurrences
         filters.push(ngsimFilters.firstKeeper());

### 2015-02-01 project1/data_processing/ngsimFilters.js
Adding a filter for turning

@@ -34,6 +34,16 @@ var filterIntersection = function(intersection) {
 //
 
 /**
+ * Filters by turn direction
+ */
+var filterMovement = function(direction) {
+    return filterFactory(19, direction);
+};
+filterMovement.THROUGH = '1';
+filterMovement.LEFT = '2';
+filterMovement.RIGHT = '3';
+
+/**
  * Factory for a filter that only keeps the first occurrence of car
  */
 var keepFirst = function() {
@@ -56,6 +66,7 @@ var keepFirst = function() {
 module.exports = {
     direction: filterDirection,
     intersection: filterIntersection,
+    movement: filterMovement,
     firstKeeper: keepFirst,
     factory: filterFactory
 };

### 2015-02-01 project1/data_processing/gulpfile.js
Trying out new filter

@@ -4,7 +4,8 @@ var inputDistribution = require('./inputDistributionTaskFactory.js');
 
 gulp.task('northbound', inputDistribution({
     direction: ngsim.direction.NORTH,
-    intersection: '3'
+    intersection: '3',
+    movement: ngsim.movement.THROUGH
 }));
 
 gulp.task('southbound', inputDistribution({

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js
Making task helper more robust to new filters

@@ -5,6 +5,17 @@ var mean = require('./mean.js');
 var performGoodnessOfFit = require('./performGoodnessOfFit.js');
 var histogram = require('./histogram.js');
 
+var getFilters = function(options) {
+    var filters = [];
+
+    // loop over all of the config options provided
+    for(var key in options) {
+        filters.push(ngsimFilters[key](options[key]));
+    }
+
+    return filters;
+};
+
 /**
  * Factory function for building a gulp task for filtering the NGSIM data for a certain direction and intersection of the road
  */
@@ -12,17 +23,7 @@ module.exports = function(options) {
     return function(taskDone) {
         var startTimes = [];
 
-        var filters = [];
-
-        // Only one direction
-        if(options.direction) {
-            filters.push(ngsimFilters.direction(options.direction));
-        }
-
-        // only through our intersection
-        if(options.intersection) {
-            filters.push(ngsimFilters.intersection(options.intersection));
-        }
+        var filters = getFilters(options);
 
         // only keep the first occurrences
         filters.push(ngsimFilters.firstKeeper());

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js
Adding a comment

@@ -5,6 +5,9 @@ var mean = require('./mean.js');
 var performGoodnessOfFit = require('./performGoodnessOfFit.js');
 var histogram = require('./histogram.js');
 
+/**
+ * Helper for appending all filters to the csv file
+ */
 var getFilters = function(options) {
     var filters = [];
 

### 2015-02-01 project1/data_processing/ngsimFilters.js
New filters

@@ -44,6 +44,21 @@ filterMovement.LEFT = '2';
 filterMovement.RIGHT = '3';
 
 /**
+ * Filters by origin
+ */
+var filterOrigin = function(origin) {
+    return filterFactory(14, origin);
+};
+// TODO: Name the origins
+
+/**
+ * Filters by destination
+ */
+var filterDestination = function(destination) {
+    return filterFactory(15, destination);
+};
+
+/**
  * Factory for a filter that only keeps the first occurrence of car
  */
 var keepFirst = function() {
@@ -67,6 +82,8 @@ module.exports = {
     direction: filterDirection,
     intersection: filterIntersection,
     movement: filterMovement,
+    origin: filterOrigin,
+    destination: filterDestination,
     firstKeeper: keepFirst,
     factory: filterFactory
 };

### 2015-02-01 project1/data_processing/gulpfile.js
Trying a filter for platoon leader?

@@ -2,10 +2,15 @@ var gulp = require('gulp');
 var ngsim = require('./ngsimFilters.js');
 var inputDistribution = require('./inputDistributionTaskFactory.js');
 
-gulp.task('northbound', inputDistribution({
+gulp.task('northboundPlatoonLeaders', inputDistribution({
     direction: ngsim.direction.NORTH,
     intersection: '3',
-    movement: ngsim.movement.THROUGH
+    isLeader: true
+}));
+
+gulp.task('northbound', inputDistribution({
+    direction: ngsim.direction.NORTH,
+    intersection: '3'
 }));
 
 gulp.task('southbound', inputDistribution({
@@ -13,4 +18,4 @@ gulp.task('southbound', inputDistribution({
     intersection: '3'
 }));
 
-gulp.task('default', ['northbound']);
+gulp.task('default', ['northboundPlatoonLeaders', 'northbound']);

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js

@@ -38,6 +38,7 @@ module.exports = function(options) {
         });
 
         var finish = function() {
+
             // calculate the intervals between arrivals (in seconds)
             var startOffsets = array_interval(startTimes).map(function(interval) {
                 return interval / 1000;

### 2015-02-01 project1/data_processing/ngsimFilters.js

@@ -58,6 +58,10 @@ var filterDestination = function(destination) {
     return filterFactory(15, destination);
 };
 
+var filterIsLeader = function() {
+    return filterFactory(20, '0');
+};
+
 /**
  * Factory for a filter that only keeps the first occurrence of car
  */
@@ -84,6 +88,7 @@ module.exports = {
     movement: filterMovement,
     origin: filterOrigin,
     destination: filterDestination,
+    isLeader: filterIsLeader,
     firstKeeper: keepFirst,
     factory: filterFactory
 };

### 2015-02-01 project1/data_processing/csvFilter.js
Back to basics

@@ -1,31 +1,0 @@
-var csv = require('csv');
-var fs = require('fs');
-
-/**
- * Helper method for querying the CSV file
- */
-var getCsv = function() {
-    // return a stream filtered and transformed into a new format
-    return fs.createReadStream('data.csv')
-        .pipe(csv.parse());
-};
-
-/**
- * Provides an easy way to hook into filtering the CSV file
- */
-module.exports = function(filters, finishCode) {
-
-    var filterCount = filters.length;
-
-    var stream = getCsv();
-
-    // tag on all of the filters
-    for(var i = 0; i < filterCount; i++) {
-        stream = stream.pipe(csv.transform(filters[i]));
-    }
-
-    // now add the end stream listener
-    stream.on('finish', finishCode);
-
-    return stream;
-};

### 2015-02-01 project1/data_processing/gulpfile.js

@@ -1,21 +1,5 @@
 var gulp = require('gulp');
-var ngsim = require('./ngsimFilters.js');
-var inputDistribution = require('./inputDistributionTaskFactory.js');
 
-gulp.task('northboundPlatoonLeaders', inputDistribution({
-    direction: ngsim.direction.NORTH,
-    intersection: '3',
-    isLeader: true
-}));
-
-gulp.task('northbound', inputDistribution({
-    direction: ngsim.direction.NORTH,
-    intersection: '3'
-}));
-
-gulp.task('southbound', inputDistribution({
-    direction: ngsim.direction.SOUTH,
-    intersection: '3'
-}));
-
-gulp.task('default', ['northboundPlatoonLeaders', 'northbound']);
+gulp.task('default', function(cb) {
+    cb();
+});

### 2015-02-01 project1/data_processing/inputDistributionTaskFactory.js

@@ -1,57 +1,0 @@
-var csvFilter = require('./csvFilter.js');
-var ngsimFilters = require('./ngsimFilters.js');
-var array_interval = require('./array_interval.js');
-var mean = require('./mean.js');
-var performGoodnessOfFit = require('./performGoodnessOfFit.js');
-var histogram = require('./histogram.js');
-
-/**
- * Helper for appending all filters to the csv file
- */
-var getFilters = function(options) {
-    var filters = [];
-
-    // loop over all of the config options provided
-    for(var key in options) {
-        filters.push(ngsimFilters[key](options[key]));
-    }
-
-    return filters;
-};
-
-/**
- * Factory function for building a gulp task for filtering the NGSIM data for a certain direction and intersection of the road
- */
-module.exports = function(options) {
-    return function(taskDone) {
-        var startTimes = [];
-
-        var filters = getFilters(options);
-
-        // only keep the first occurrences
-        filters.push(ngsimFilters.firstKeeper());
-
-        // track the start times
-        filters.push(function(row) {
-            startTimes.push(Number(row[3]));
-            return row;
-        });
-
-        var finish = function() {
-
-            // calculate the intervals between arrivals (in seconds)
-            var startOffsets = array_interval(startTimes).map(function(interval) {
-                return interval / 1000;
-            });
-
-            var chiSquaredResults = performGoodnessOfFit(startOffsets, 25);
-            console.log('N = ' + startOffsets.length);
-            console.log("Chi-Squared Test : p = " + chiSquaredResults.probability + " chi^2 = " + chiSquaredResults.chiSquared);
-            console.log(histogram(startOffsets, { }));
-
-            taskDone();
-        };
-
-        csvFilter(filters, finish);
-    };
-};

### 2015-02-01 project1/data_processing/mean.js

@@ -1,8 +1,0 @@
-/**
- * Calculates the average of a list.
- */
-module.exports = function(items) {
-    return items.reduce(function(accumulator, item) {
-        return accumulator + item;
-    }, 0) / items.length;
-};

### 2015-02-01 project1/data_processing/ngsimFilters.js

@@ -1,94 +1,0 @@
-/**
- * Basic function for building a filter for a given row on a certain column
- */
-var filterFactory = function(columnIndex, value) {
-    return function(row) {
-        if(row[columnIndex] != value) {
-            return;
-        }
-
-        return row;
-    };
-};
-
-/**
- * Factory for building a filter for direction
- */
-var filterDirection = function(direction) {
-    return filterFactory(18, direction); // 18 is direction
-};
-
-filterDirection.EAST = '1';
-filterDirection.NORTH = '2';
-filterDirection.WEST = '3';
-filterDirection.SOUTH = '4';
-
-/**
- * Filters by intersection
- */
-var filterIntersection = function(intersection) {
-    return filterFactory(16, intersection);
-};
-
-// TODO: Name the intersections, so users have something to work with
-//
-
-/**
- * Filters by turn direction
- */
-var filterMovement = function(direction) {
-    return filterFactory(19, direction);
-};
-filterMovement.THROUGH = '1';
-filterMovement.LEFT = '2';
-filterMovement.RIGHT = '3';
-
-/**
- * Filters by origin
- */
-var filterOrigin = function(origin) {
-    return filterFactory(14, origin);
-};
-// TODO: Name the origins
-
-/**
- * Filters by destination
- */
-var filterDestination = function(destination) {
-    return filterFactory(15, destination);
-};
-
-var filterIsLeader = function() {
-    return filterFactory(20, '0');
-};
-
-/**
- * Factory for a filter that only keeps the first occurrence of car
- */
-var keepFirst = function() {
-    var seenCars = {};
-    var firstTracker = function(row) {
-        var car_id = row[0];
-
-        // if we've seen the car skip it
-        if(seenCars[car_id]) {
-            return null;
-        }
-
-        seenCars[car_id] = true;
-        return row;
-    };
-
-    return firstTracker;
-};
-
-module.exports = {
-    direction: filterDirection,
-    intersection: filterIntersection,
-    movement: filterMovement,
-    origin: filterOrigin,
-    destination: filterDestination,
-    isLeader: filterIsLeader,
-    firstKeeper: keepFirst,
-    factory: filterFactory
-};

### 2015-02-02 project1/data_processing/gulpfile.js
Adding some outlier detection

@@ -1,5 +1,29 @@
 var gulp = require('gulp');
+var fs = require('fs');
+var array_interval = require('./array_interval.js');
+var histogram = require('./histogram.js');
+var onevar = require('./one_var_stats.js');
+
+var readIntervals = function(path) {
+    var timestamps = fs.readFileSync(path)
+        .toString()
+        .split('\n')
+        .filter(function(item) {
+            return item.length > 0;
+        })
+        .map(function(item) {
+            return Number(item) / 1000;
+        })
+        .sort(function(a, b) {
+            return a - b;
+        });
+
+    return array_interval(timestamps);
+};
 
 gulp.task('default', function(cb) {
+    var intervals = readIntervals('data/all_northbound.csv');
+    var inliers = onevar.inliers(intervals);
+    console.log(histogram(inliers));
     cb();
 });

### 2015-02-02 project1/data_processing/one_var_stats.js

@@ -1,0 +1,79 @@
+/**
+ * Calculates the median of a set of numbers
+ */
+var median = function(numbers) {
+    var sorted = numbers.sort(function(a, b) {
+        return Number(a) - Number(b);
+    });
+
+    var middle = Math.floor(sorted.length / 2);
+    if(sorted.length % 2 == 1) {
+        return sorted[middle];
+    }
+    else {
+        return 0.5 * (sorted[middle] + sorted[middle-1]);
+    }
+};
+
+/**
+ * Calculates the first quartile of the set of numbers
+ */
+var firstQuartile = function(numbers) {
+    var medianOfSet = median(numbers);
+    var lowerHalf = numbers.filter(function(item) {
+        return item < medianOfSet;
+    });
+
+    return median(lowerHalf);
+};
+
+/**
+ * Calculates the third quartile of the set of numbers
+ */
+var thirdQuartile = function(numbers) {
+    var medianOfSet = median(numbers);
+    var upperHalf = numbers.filter(function(item) {
+        return item > medianOfSet;
+    });
+
+    return median(upperHalf);
+};
+
+/**
+ * Calculates the interquartile range
+ */
+var IQR = function(numbers) {
+    var q1 = firstQuartile(numbers);
+    var q3 = thirdQuartile(numbers);
+    return q3 - q1;
+};
+
+/**
+ * Gets the outliers of a given set
+ */
+var outliers = function(numbers) {
+    var medianOfSet = median(numbers);
+    var iqrOfSet = IQR(numbers);
+    return numbers.filter(function(item) {
+        return Math.abs(item - medianOfSet) >= iqrOfSet;
+    });
+};
+
+/**
+ * Gets the non-outliers of a given set
+ */
+var inliers = function(numbers) {
+    var medianOfSet = median(numbers);
+    var iqrOfSet = IQR(numbers);
+    return numbers.filter(function(item) {
+        return Math.abs(item - medianOfSet) <= iqrOfSet;
+    });
+};
+
+module.exports = {
+    median: median,
+    firstQuartile: firstQuartile,
+    thirdQuartile: thirdQuartile,
+    outliers: outliers,
+    inliers: inliers
+};

### 2015-02-02 project1/data_processing/gulpfile.js
Adding in histograms of outliers

@@ -21,9 +21,18 @@ var readIntervals = function(path) {
     return array_interval(timestamps);
 };
 
-gulp.task('default', function(cb) {
+gulp.task('northbound-inliers', function(cb) {
     var intervals = readIntervals('data/all_northbound.csv');
     var inliers = onevar.inliers(intervals);
     console.log(histogram(inliers));
     cb();
 });
+
+gulp.task('northbound-outliers', function(cb) {
+    var intervals = readIntervals('data/all_northbound.csv');
+    var outliers = onevar.outliers(intervals);
+    console.log(histogram(outliers));
+    cb();
+});
+
+gulp.task('default', ['northbound-inliers', 'northbound-outliers']);

### 2015-02-03 project1/data_processing/expectedExponentialFrequencies.js
Trying out chi-squared fitting. Somewhat better...

@@ -1,27 +1,0 @@
-var exponential = require('./exponential.js');
-
-/**
- * calculates the integral of the exponential pdf between two values
- */
-var probabilityBetween = function(a, b, lambda) {
-    return exponential.cdf(b, lambda) - exponential.cdf(a, lambda);
-};
-
-/**
- * Calculates the expected probabilities for a set of bins provided by:
- *
- * @param binWidth The width of each bin
- * @param binCount The number of bins
- * @param lambda   The lambda to use
- */
-module.exports = function(binWidth, binCount, lambda) {
-    var frequencies  = [];
-
-    for(var i = 0; i < binCount; i++) {
-        var leftBoundary = i * binWidth;
-        var rightBoundary = (i + 1) * binWidth;
-        frequencies.push(probabilityBetween(leftBoundary, rightBoundary, lambda));
-    }
-
-    return frequencies;
-};

### 2015-02-03 project1/data_processing/gulpfile.js

@@ -3,6 +3,7 @@ var fs = require('fs');
 var array_interval = require('./array_interval.js');
 var histogram = require('./histogram.js');
 var onevar = require('./one_var_stats.js');
+var goodnessOfFit = require('./performGoodnessOfFit.js');
 
 var readIntervals = function(path) {
     var timestamps = fs.readFileSync(path)
@@ -25,6 +26,10 @@ gulp.task('northbound-inliers', function(cb) {
     var intervals = readIntervals('data/all_northbound.csv');
     var inliers = onevar.inliers(intervals);
     console.log(histogram(inliers));
+
+    var chiSquaredResults = goodnessOfFit(inliers, 25);
+    console.log(chiSquaredResults.chiSquared);
+    console.log(chiSquaredResults.probability);
     cb();
 });
 

### 2015-02-03 project1/data_processing/performGoodnessOfFit.js

@@ -1,22 +1,36 @@
-var expectedExponentialFrequencies = require('./expectedExponentialFrequencies.js');
-var mean = require('./mean.js');
 var Bin = require('./bin.js');
+var gammaCDF = require('gamma-distribution').cdf;
+var fit = require('gamma-distribution').fit;
 var chiSquaredTest = require('chi-squared-test');
+
+var gammaCdfRange = function(a, b, k, theta) {
+    return gammaCDF(b, k, theta) - gammaCDF(a, k, theta);
+};
+
+var calculateExpectations = function(binWidth, binCount, k, theta) {
+    var bins = [];
+    for(var i = 0; i < binCount; i++) {
+        bins.push( gammaCdfRange(binWidth * i, binWidth * (i + 1), k, theta) );
+    }
+
+    return bins;
+};
 
 /**
  * Performs an exponential goodness-of-fit test to determine how good a fit some raw data is. Essentially performs a
  * chi-squared test
  */
 module.exports = function(rawData, binCount) {
+    rawData = rawData.filter(function(item) {
+        return item > 0;
+    });
     var binnedData = new Bin(rawData, binCount);
-
-    // The maximum likelihood lambda value 1 / mean
-    var mleLambda = 1 / mean(rawData);
 
     var N = rawData.length;
 
     // the expected frequencies
-    var expectations = expectedExponentialFrequencies(binnedData.binWidth, binCount, mleLambda).map(function(percent) {
+    var params = fit(rawData);
+    var expectations = calculateExpectations(binnedData.binWidth, binCount, params.k, params.theta).map(function(percent) {
         return percent * N;
     });
 

### 2015-02-03 project1/data_processing/em.js
Stubbing out an expectation maximization algorithm

@@ -1,0 +1,112 @@
+/**
+ * Finds the likelihood for a given pdf, x value, and parameter list
+ *
+ * @param pdf        The pdf function to use
+ * @param x          The x value
+ * @param parameters The PDF parameters
+ */
+var likelihood = function(pdf, x, parameters) {
+    // copy the array
+    var pdfParameters = parameters.slice();
+
+    // push x on the front
+    pdfParameters.unshift(x);
+
+    // call the pdf
+    return pdf.apply(pdf, pdfParameters);
+};
+
+/**
+ * Chooses the most likely distribution that the given x belongs to
+ *
+ * @param pdf           The pdf function to use
+ * @param x             The x value to test
+ * @param parameterList The list of parameters to try
+ */
+var mostLiklyDistribution = function(pdf, x, parameterList) {
+    var N = parameterList.length;
+    var greatestLikelihood = -1;
+    var greatestLikelihoodIndex = 0;
+
+    for(var i = 0; i < N; i++) {
+        var currentLikelihood = likelihood(pdf, x, parameterList[i]);
+
+        if(currentLikelihood > greatestLikelihood) {
+            greatestLikelihood = currentLikelihood;
+            greatestLikelihoodIndex = i;
+        }
+    }
+
+    return greatestLikelihoodIndex;
+};
+
+/**
+ * Performs distribution assignment, given a pdf, set of x values, and parameter list
+ *
+ * @param pdf           The pdf to use
+ * @param xList         A list of x values to assign
+ * @param parameterList A list of parameters for the mixture model distribution
+ */
+var assignToDistribution = function(pdf, xList, parameterList) {
+    var distributionCount = parameterList.length;
+    var i;
+
+    // initialize assignments
+    var assignments = [];
+    for(i = 0; i < distributionCount; i++) {
+        assignments.push([]);
+    }
+
+    // now loop over each x and assign accordingly
+    var xCount = xList.length;
+    for(i = 0; i < xCount; i++) {
+        var index = mostLiklyDistribution(pdf, xList[i], parameterList);
+        assignments[index].push(xList[i]);
+    }
+
+    return assignments;
+};
+
+/**
+ * Refits data to a distribution for each assignment in the list
+ *
+ * @param fitFunction The function to use for fitting the data
+ * @param assignments The list of current assignments
+ */
+var refitData = function(fitFunction, assignments) {
+    var distributionCount = assignments.length;
+    var parameterList = [];
+
[…]

### 2015-02-03 project1/data_processing/gulpfile.js
Making the output a little nicer

@@ -4,6 +4,7 @@ var array_interval = require('./array_interval.js');
 var histogram = require('./histogram.js');
 var onevar = require('./one_var_stats.js');
 var goodnessOfFit = require('./performGoodnessOfFit.js');
+var fit = require('gamma-distribution').fit;
 
 var readIntervals = function(path) {
     var timestamps = fs.readFileSync(path)
@@ -24,12 +25,18 @@ var readIntervals = function(path) {
 
 gulp.task('northbound-inliers', function(cb) {
     var intervals = readIntervals('data/all_northbound.csv');
-    var inliers = onevar.inliers(intervals);
-    console.log(histogram(inliers));
+    intervals = intervals.filter(function(item) {
+        return item > 0.001;
+    });
+    console.log(histogram(intervals));
 
-    var chiSquaredResults = goodnessOfFit(inliers, 25);
-    console.log(chiSquaredResults.chiSquared);
-    console.log(chiSquaredResults.probability);
+    console.log('Single Gamma Chi-Squared Fit');
+    var chiSquaredResults = goodnessOfFit(intervals, 25);
+    var fittedValues = fit(intervals);
+    console.log('k = ' + fittedValues.k + ' theta = ' + fittedValues.theta);
+    console.log('chi-squared = ' + chiSquaredResults.chiSquared);
+    console.log('p = ' + chiSquaredResults.probability);
+
     cb();
 });
 

### 2015-02-03 project1/data_processing/em.js
Forgot to provide the parameter

@@ -94,7 +94,7 @@ var refitData = function(fitFunction, assignments) {
  *                          also how the al
  * @param xList             The list of data to fit
  */
-module.exports = function(pdf, fitData, initialParameters) {
+module.exports = function(pdf, fitData, initialParameters, xList) {
     var maxSteps = 1000;
 
     var assignments;

### 2015-02-03 project1/data_processing/gulpfile.js
That junk worked the first time?

@@ -4,6 +4,9 @@ var array_interval = require('./array_interval.js');
 var histogram = require('./histogram.js');
 var onevar = require('./one_var_stats.js');
 var goodnessOfFit = require('./performGoodnessOfFit.js');
+
+var gammaPdf = require('gamma-distribution').pdf;
+var em = require('./em.js');
 var fit = require('gamma-distribution').fit;
 
 var readIntervals = function(path) {
@@ -36,6 +39,20 @@ gulp.task('northbound-inliers', function(cb) {
     console.log('k = ' + fittedValues.k + ' theta = ' + fittedValues.theta);
     console.log('chi-squared = ' + chiSquaredResults.chiSquared);
     console.log('p = ' + chiSquaredResults.probability);
+
+    console.log('Mixture Model Chi-Squared Fit');
+    var emFit = function(data) {
+        var result = fit(data);
+        return [result.k, result.theta]; // since EM wants an array
+    };
+
+    var initialParams = [
+        [1, 2],
+        [7.5, 1],
+    ];
+
+    var result = em(gammaPdf, emFit, initialParams, intervals);
+    console.log(result);
 
     cb();
 });

### 2015-02-03 project1/data_processing/expectedFrequenciesForMixtureModel.js
Adding some helpers for testing mixture model fits

@@ -1,0 +1,88 @@
+/**
+ * Calculates the weights for each probability distribution, based on how many points are assigned to it
+ *
+ * @param assignments A jagged array of assignments
+ * @return array
+ */
+var calculateMixtureProbabilities = function(assignments) {
+    var mixtureProbabilities = [];
+    var distributionCount = assignments.length;
+
+    // add the counts in
+    for(var i = 0; i < distributionCount; i++) {
+        mixtureProbabilities.push(assignments[i].length);
+    }
+
+    // get totals
+    var totalPoints = mixtureProbabilities.reduce(function(accumulator, count) {
+        return accumulator + count;
+    }, 0);
+
+    // scale down to probabilities
+    return mixtureProbabilities.map(function(count) {
+        return count / totalPoints;
+    });
+};
+
+/**
+ * Uses a cdf to calculate the area between two x-values
+ *
+ * @param cdf        The cumulative distribution function
+ * @param a          The left limit of integration
+ * @param b          The right limit of integration
+ * @param parameters The parameters to use for the distribution
+ */
+var cdfBetween = function(cdf, a, b, parameters) {
+    var aParameters = parameters.slice();
+    aParameters.unshift(a);
+    var bParameters = parameters.slice();
+    bParameters.unshift(b);
+
+    return cdf.apply(cdf, bParameters) - cdf.apply(cdf, aParameters);
+};
+
+/**
+ * Calculates the full probability between two points for the true mixture model
+ *
+ * @param cdf                  The cumulative distribution function to use
+ * @param a                    The left limit of integration
+ * @param b                    The right limit of integration
+ * @param parameterList        The list of parameters to use for each distribution
+ * @param mixtureProbabilities The mixture probabilites for the mixture model
+ *
+ * @return Number
+ */
+var calculateMixtureCdfBetween = function(cdf, a, b, parameterList, mixtureProbabilities) {
+    var total = 0;
+    var distributionCount = mixtureProbabilities.length;
+
+    for(var i = 0; i < distributionCount; i++) {
+        total += mixtureProbabilities[i] * cdfBetween(cdf, a, b, parameterList[i]);
+    }
+
+    return total;
+};
+
+/**
+ * Builds an array of the expected probabilities to fall within a set of buckets
+ *
+ * @param binWidth      the width of each bin
+ * @param binCount      The number of bins to calculate for
+ * @param cdf           The cumulative distribution function to use
[…]

### 2015-02-03 project1/data_processing/gulpfile.js

@@ -1,13 +1,15 @@
 var gulp = require('gulp');
+var Bin = require('./bin.js');
 var fs = require('fs');
 var array_interval = require('./array_interval.js');
 var histogram = require('./histogram.js');
 var onevar = require('./one_var_stats.js');
 var goodnessOfFit = require('./performGoodnessOfFit.js');
 
-var gammaPdf = require('gamma-distribution').pdf;
+var gamma = require('gamma-distribution');
 var em = require('./em.js');
-var fit = require('gamma-distribution').fit;
+var expectedFrequenciesForMixtureModel = require('./expectedFrequenciesForMixtureModel.js');
+var chiSquaredTest = require('chi-squared-test');
 
 var readIntervals = function(path) {
     var timestamps = fs.readFileSync(path)
@@ -31,29 +33,38 @@ gulp.task('northbound-inliers', function(cb) {
     intervals = intervals.filter(function(item) {
         return item > 0.001;
     });
+    intervals = onevar.inliers(intervals);
     console.log(histogram(intervals));
 
     console.log('Single Gamma Chi-Squared Fit');
     var chiSquaredResults = goodnessOfFit(intervals, 25);
-    var fittedValues = fit(intervals);
+    var fittedValues = gamma.fit(intervals);
     console.log('k = ' + fittedValues.k + ' theta = ' + fittedValues.theta);
     console.log('chi-squared = ' + chiSquaredResults.chiSquared);
     console.log('p = ' + chiSquaredResults.probability);
 
     console.log('Mixture Model Chi-Squared Fit');
     var emFit = function(data) {
-        var result = fit(data);
+        var result = gamma.fit(data);
         return [result.k, result.theta]; // since EM wants an array
     };
 
     var initialParams = [
-        [1, 2],
-        [7.5, 1],
+        [1, 5],
+        [2.5, 2.5],
     ];
 
-    var result = em(gammaPdf, emFit, initialParams, intervals);
-    console.log(result);
+    var result = em(gamma.pdf, emFit, initialParams, intervals);
 
+    for(var i = 0; i < result.assignments.length; i++) {
+        var chiSquaredResults = goodnessOfFit(result.assignments[i], 25);
+        var fittedValues = gamma.fit(result.assignments[i]);
+        console.log('k = ' + fittedValues.k + ' theta = ' + fittedValues.theta);
+        console.log('chi-squared = ' + chiSquaredResults.chiSquared);
+        console.log('p = ' + chiSquaredResults.probability);
+    }
+
+    
     cb();
 });
 

### 2015-02-03 project1/checkpoint/main.tex
Fleshing out the conceptual model
Fleshing out conceptual model some more

@@
 conceptual version of this model from which we can plan our implementation. Data from \cite{cts12} was used extensively % TODO: Put this somewhere else
-for modeling purposes in this project. Our first step will be to identify the entities present in our model, and then consider their interactions (or the
-inherent ``activies'') within the simulation.
+for modeling purposes in this project. Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
+we'll utilize the publisher-subscriber pattern (see \cite{pubsub}) to promote better decoupling.
+Given this architecture, our first step will be to identify the entities present in our model, and then consider their
+interactions (or the inherent ``activies'') within the simulation. This will allow us to then design the events that
+are important
 
+\subsection{Entities}
 The first entity, unsurprisingly, is a single car. Our representation of a car is clearly less detailed than the real
@@
 
+\subsection{Basic Interactions}
 From the intersection entity, we can glean an important interaction that occurs: The changing of light signal. When a
 signal changes, each car in-queue will be forced to change it's state, whether it's waiting or currently in motion.
+From an event-driven perspective, this means that any listener would need to be able to know \emph{how} the light
+changed in order to act accordingly.
+
+An intersection serves as holding point for cars traveling along the road so if a light is red, cars are forced to
+queue up at the intersection until green. This lends itself to a queue-style data structure that encorporates both
+position and acceleration. Car acceleration is a complex topic, but models have been constructed \cite{bonneson}
+\cite{herman_et_al}. These models are generally continuous, and can generally be represented as a differential equation
+due to the relationships between acceleration and velocity, as can be seen in \cite{briggs} \cite{deceleration} as well.
+Instead, we will
+simply consider the various events that trigger changes in behavior. For instance, if a car slows down, the car behind
+it will be notified and change it's driving patterns accordingly.
+
+\subsection{Events}
+Based on our observations from above, we can now construct a set of events that are pertinent to this application.
+
+First off, signal changes consititute the major event in the system. When a signal changes, the event must contain
+\emph{when} the signal changed, and \emph{what} the signal changed to. The light itself will listen for this event
+and will enqueue the next signal change based on the signal change time. These signal changes could be enqueued
+completely upfront, but allowing this to be done on the fly allows the user to tweak these values during the simulation.
+
+The car model will listen for this event as well. When a currently traveling car sees a red light, it will calculate
+\emph{where} it needs to stop, and \emph{when} it will come to a stop. This will be done using a deceleration model
+similar to \cite{deceleration}.
 

### 2015-02-03 project1/checkpoint/main.tex
Finishing up conceptual model
Adding a little more text

@@
 \emph{where} it needs to stop, and \emph{when} it will come to a stop. This will be done using a deceleration model
-similar to \cite{deceleration}.
+similar to \cite{deceleration}. If the light turns to green, the car instead needs to start accelerating (using one of
+the models mentioned above). The model will adapt to light changes during movement by recalculating its estimated
+exit time of the system.
 
@@
 We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
-consider the car arrival time, and car length when modeling cars. Due to the discrete nature of the application, we'll
-ignore some of the more "continuous" aspects, such car velocity and acceleration and encompass them into calculation
-of event times.
+consider basic aspects of a car: when it enters the system, and it's intended destination (turning left or right at
+12th street or going straight). We will assume that all cars have the same speed entering the system, and that they
+have the same length. These values can be calculated from the NGSIM data set \cite{ngsim}.
+
+The distribution of car arrival is very important, but has been difficult to model.
 

### 2015-02-03 project1/checkpoint/gulpfile.js
Adding figures to the file

@@ -16,15 +16,13 @@ gulp.task('clean', function(cb) {
 
 
 gulp.task('default', ['clean'], function(cb) {
-    var latex = 'latex -halt-on-error main';
+    var latex = 'pdflatex -halt-on-error main';
     var bibtex = 'bibtex main';
-    var dvipdf = 'dvipdf main.dvi main.pdf';
     
     async.series([
         asyncCommand(latex),
         asyncCommand(bibtex),
         asyncCommand(latex),
-        asyncCommand(latex),
-        asyncCommand(dvipdf)
+        asyncCommand(latex)
     ], cb);
 });

### 2015-02-03 project1/checkpoint/main.tex
Adding figures to the file
Finishing up the assumptions section

@@
 
-\usepackage{url, hyperref}
+\usepackage{url, hyperref, graphicx}
 
@@
 
-The distribution of car arrival is very important, but has been difficult to model.
+\begin{figure}
+\begin{center}  
+\includegraphics[width=4in]{../northbound.png}  
+\caption{\small \sl Northbound Arrival Intervals \label{fig:northbound_arrival}}  
+\end{center}  
+\end{figure}
+
+\begin{figure}  
+\begin{center}
+\includegraphics[width=4in]{../southbound.png}  
+\caption{\small \sl Southbound Arrival Intervals \label{fig:southbound_arrival}}  
+\end{center}  
+\end{figure}
+
+The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
+northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
+Model fitting was attempted for both $\Gamma$ and $\Gamma$ mixture models on the raw data, but resulted in poor
+fits. Furthermore, pruning outliers via the common interquartile range approach, where values are rejected when
+they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.
+
+So, instead we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
+and simply use the observed frequency distribution as our input distribution. As the project progresses we can
+refine this model.
+
+The current light timing of the system is known, and provided in \cite{ngsim}, but we'll allow the user to set
+those values dynamically via a UI.
 
 \subsection{Simulation Output}
+The most important statistic we'll want to collect during the simulation is the thoughput of the simulation. We'll
+want to observe how the light timing affects the average travel time for cars. It will also to see how variable this
+is. For instance, do cars travelling on Peachtree St. have similar travel times, despite their arrival time, or does
+their arrival drastically lengthen their time to travel the stretch of road we are modeling.
 
@@
 
+Another simplification to consider is lane changing. Lane changing models do exist \cite{lanechanging} \cite{lanechanging2},
+but can be complex, requiring modeling of driver choices and gap acceptance for drivers. Peachtree is a two lane road,
+so lane changes are possible. However, we will make the assumption that drivers by this point have already made any
+discretionary lane changes and traffic is evened out between lanes. Also, we will assume left turning drivers will be
+in the left lane already. Similarly, right turning drivers will already be in the right lane.
+
 \section{Simulation Software Architecture}

### 2015-02-03 project1/checkpoint/main.tex
Talking about the data analysis tools

@@
 
-So, instead we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
+Given the difficulties with modelling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
 and simply use the observed frequency distribution as our input distribution. As the project progresses we can
@@
 those values dynamically via a UI.
+
+\subsubsection{Data Analysis}
+In the \texttt{data\_processing} folder, most of code used for distribution fitting and statistics can be found there.
+The data analysis code was written in Javascript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
+to manage dependencies and those dependencies can be seen in the \texttt{package.json} file. Of those packages, I have
+authored the \texttt{chi-squared-test} module \cite{chiSquaredNpm}, and the \texttt{gamma-distribution}
+\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used a build tool for automating tasks (including building this paper).
 
@@
 \section{Simulation Software Architecture}
-Simulation Software
+The simulation software attempts to utilize
 

### 2015-02-03 project1/checkpoint/main.tex
Adding the software simulation section

@@
 \section{Simulation Software Architecture}
-The simulation software attempts to utilize
+Because of the event-driven architecture, the system can be though of as a set of event emitters and event listeners.
+In some cases, an object may both emit and listen for events.
+
+The first event emitter to consider is the \texttt{CarEmitter}. The \texttt{CarEmitter} uses the measured input
+distribution we have for cars and generates a set of cars randomly. It then emits these cars as individual
+\texttt{car:arrived} event. There will generally be two \texttt{CarEmitter}s, one for the North and South direction.
+
+The second emitter will be the \texttt{LightSignal}. It will emit \texttt{light:changed} events, passing the light color.
+Also, instead of enqueueing these events upfront, it will also listen for \texttt{light:changed} so that its
+changing duration can modified on the fly (using the UI).
+
+The \texttt{IntersectionQueue} holds a list of cars lined up at the intersection. It also maintains the
+current signal state, so that it can send cars directly through the intersection if the light is green. It responds
+to pretty much every event mentioned above. When the light changes, it will need update it's current signal state.
+This may mean sending some cars off of the queue and stop holding cars, or to no longer let cars through and make them
+wait. Furthermore, when a car arrives it will handle that event by deciding whether the car can pass through or needs
+to wait on the light. Also, this class emits events. Whenever a car finally crosses the intersection, it emits
+a \texttt{car:exited} event.
+
+Since we want to gather metrics regarding the throughput of cars, we'll also provide a listener class
+\texttt{ThroughputStatistics} that will listen for cars to complete their journey through the system. This will
+be reported in the UI in real time.
 

### 2015-02-04 project1/simulation/gulpfile.js
Stubbing out the actual application

@@ -1,0 +1,14 @@
+var gulp = require('gulp');
+var coffeeify = require('gulp-coffeeify');
+
+gulp.task('bundle', function() {
+    return gulp.src('coffee/main.coffee')
+        .pipe(coffeeify())
+        .pipe(gulp.dest('.'));
+});
+
+gulp.task('default', ['bundle']);
+
+gulp.task('watch', ['bundle'], function() {
+    gulp.watch('coffee/*', 'bundle');
+});

### 2015-02-04 project1/simulation/gulpfile.js
Adding a clean task

@@ -1,10 +1,15 @@
 var gulp = require('gulp');
+var del = require('del');
 var coffeeify = require('gulp-coffeeify');
 
 gulp.task('bundle', function() {
     return gulp.src('coffee/main.coffee')
         .pipe(coffeeify())
         .pipe(gulp.dest('.'));
+});
+
+gulp.task('clean', function(cb) {
+    del('main.js', cb);
 });
 
 gulp.task('default', ['bundle']);

### 2015-02-04 project1/simulation/gulpfile.js
Adding in uglify

@@ -1,10 +1,12 @@
 var gulp = require('gulp');
 var del = require('del');
 var coffeeify = require('gulp-coffeeify');
+var uglify = require('gulp-uglify');
 
 gulp.task('bundle', function() {
     return gulp.src('coffee/main.coffee')
         .pipe(coffeeify())
+        .pipe(uglify())
         .pipe(gulp.dest('.'));
 });
 

### 2015-02-04 project1/simulation/gulpfile.js
Ignoring uglification for now

@@ -6,7 +6,7 @@ var uglify = require('gulp-uglify');
 gulp.task('bundle', function() {
     return gulp.src('coffee/main.coffee')
         .pipe(coffeeify())
-        .pipe(uglify())
+        //.pipe(uglify({ "source-map" : "main.js.map" }))
         .pipe(gulp.dest('.'));
 });
 

### 2015-02-05 project1/checkpoint/main.tex
Adding in code description paragraph

@@
 
+The simulation is written in CoffeeScript \cite{coffeescript}, which is a small language that compiles to JavaScript.
+CoffeeScripts removes some of the boilerplate associated with setting up prototype chains and other object-oriented
+concepts in JavaScript.
+Backbone \cite{backbone} along with Underscore \cite{underscore} and jQuery \cite{jquery} were used to provide some
+application structure, like basic eventing, and the sorted
+collections that are used to form the basis for the event queue and \texttt{IntersectionQueue}. Lastly, Browserify
+\cite{browserify} was used to bundle the code into a single file. Minification was done with Uglify.js \cite{uglify}.
+
 \section{Current Build}

### 2015-02-05 project1/checkpoint/main.tex
Adding a README

@@
 \section{Current Build}
-Current Build
+The simulation is currently deployed at \cite{aprilandchip} and currently runs a fixed number of cars through stepping
+at 1 second real time being equal to 100ms. Eventually, this will be adjustable in real-time, and better visualized.
 
-\cite{cts12}
+To build the application locally, you'll need to install a version of node at 0.10.0 or higher. This will be bundled
+with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
+\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
+\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
+\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
+simulation.
+
 \bibliographystyle{plain}

### 2015-02-06 project1/checkpoint/main.tex
Finishing up README
Some quick cleanup

@@
 other things. So as a result, cities must instead optimize the variables they can control. In this case, we consider
-signal light timing for the 12th St. and Peachtree Street intersection due to
-time constraints, and the sheer difficulty with modeling traffic in general between multiple intersections. In particular,
+signal light timing for the 12th St. and Peachtree Street intersection. In particular,
 we explore the effects that light timing has on traffic flow, given a source input distribution of cars.
@@
 In order to create a representative software simulation of the aforementioned system, we first begin by designing a
-conceptual version of this model from which we can plan our implementation. Data from \cite{cts12} was used extensively % TODO: Put this somewhere else
-for modeling purposes in this project. Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
+conceptual version of this model from which we can plan our implementation. We'll borrow heavily from the results and data
+in \cite{cts12} when formulating our model.
+Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
 we'll utilize the publisher-subscriber pattern (see \cite{pubsub}) to promote better decoupling.
 Given this architecture, our first step will be to identify the entities present in our model, and then consider their
-interactions (or the inherent ``activies'') within the simulation. This will allow us to then design the events that
+interactions (or the inherent ``activies'') within the simulation. This will allow us then to design the events that
 are important
@@
 \subsection{Simulation Input}
-The two main sources of input into this simulation are the cars entering the section of Peachtree St with which we are
-concerned, and the light timing of the system. The combination of these two variables are what influence the smoothness
+The two main sources of input into this simulation are the cars entering out section of Peachtree St, and the light
+timing of the system. The combination of these two variables are what influence the smoothness
 of traffic flow to the actual drivers in the system.
@@
 \includegraphics[width=4in]{../northbound.png}  
-\caption{\small \sl Northbound Arrival Intervals \label{fig:northbound_arrival}}  
+\caption{\small \sl Northbound Arrival Intervals by Order of Occurrence\label{fig:northbound_arrival}}  
 \end{center}  
@@
 \includegraphics[width=4in]{../southbound.png}
-\caption{\small \sl Southbound Arrival Intervals \label{fig:southbound_arrival}}  
+\caption{\small \sl Southbound Arrival Intervals by Order of Occurrence\label{fig:southbound_arrival}}  
 \end{center}  
@@
 northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
-Model fitting was attempted for both $\Gamma$ and $\Gamma$ mixture models on the raw data, but resulted in poor
+Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
 fits. Furthermore, pruning outliers via the common interquartile range approach, where values are rejected when
@@
 Given the difficulties with modelling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
-and simply use the observed frequency distribution as our input distribution. As the project progresses we can
+and simply plan to use the observed frequency distribution as our input distribution. As the project progresses we can
 refine this model.
@@
 In the \texttt{data\_processing} folder, most of code used for distribution fitting and statistics can be found there.
-The data analysis code was written in Javascript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
+The data analysis code was written in JavaScript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
 to manage dependencies and those dependencies can be seen in the \texttt{package.json} file. Of those packages, I have
@@
 \cite{gammaDistributionNpm}. Gulp \cite{gulp} was used a build tool for automating tasks (including building this paper).
+To run some of the data analysis code I wrote, you'll need to install and run gulp. Instructions for doing so are
+provided in the final section of this paper.
 
@@
 The simulation is written in CoffeeScript \cite{coffeescript}, which is a small language that compiles to JavaScript.
-CoffeeScripts removes some of the boilerplate associated with setting up prototype chains and other object-oriented
+CoffeeScript removes some of the boilerplate associated with setting up prototype chains and other object-oriented
 concepts in JavaScript.
@@
 
-\section{Current Build}
+\section{Current Build and Running Instructions}
 The simulation is currently deployed at \cite{aprilandchip} and currently runs a fixed number of cars through stepping
 at 1 second real time being equal to 100ms. Eventually, this will be adjustable in real-time, and better visualized.
+Cars are generated regularly at 10 (+ an uniformly random number between 0 and 1) second intervals. Light timings are
+currently hard-coded as well. Currently, we're only modeling a single direction, and ignoring turns, and variations
+between cars.
 
-To build the application locally, you'll need to install a version of node at 0.10.0 or higher. This will be bundled
+To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
 with npm. Since this project uses gulp for a task runner, you'll need to install it globally via

### 2015-02-06 project1/simulation/gulpfile.js
Adding minification back. Woops

@@ -6,7 +6,7 @@ var uglify = require('gulp-uglify');
 gulp.task('bundle', function() {
     return gulp.src('coffee/main.coffee')
         .pipe(coffeeify())
-        //.pipe(uglify({ "source-map" : "main.js.map" }))
+        .pipe(uglify({ "source-map" : "main.js.map" }))
         .pipe(gulp.dest('.'));
 });
 

### 2015-02-27 project1/checkpoint/main.tex
Basic edits in preparation for the final report

@@
 interactions (or the inherent ``activies'') within the simulation. This will allow us then to design the events that
-are important
+are important and relevant to the simulation.
 
@@
 with the variables that influence neighboring cars. For our simulation, we'll track the position and destination of the
-cars, along with the size of the car since that can influence the position of cars behind it. Also, we will track the
-current velocity of the car
+cars. Probably the most primary value associated with a car is its arrival time, since that dictates when and where it
+will fall in the queue of cars at the intersection.
 
@@
 due to the relationships between acceleration and velocity, as can be seen in \cite{briggs} \cite{deceleration} as well.
-Instead, we will
-simply consider the various events that trigger changes in behavior. For instance, if a car slows down, the car behind
-it will be notified and change it's driving patterns accordingly.
+Instead, we will simply consider the various events that trigger changes in behavior. For instance, a real driver is
+constantly observing their environment (hopefully) and making minute adjustments the entire time. However, we'll take
+a greatly simplified approach and simply precalculate an estimate of how long a car takes to traverse the section of
+road, along with a model for accelerating from a stop.
 
@@
 We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
-consider basic aspects of a car: when it enters the system, and it's intended destination (turning left or right at
-12th street or going straight). We will assume that all cars have the same speed entering the system, and that they
+consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
 have the same length. These values can be calculated from the NGSIM data set \cite{ngsim}.
@@
 Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
-fits. Furthermore, pruning outliers via the common interquartile range approach, where values are rejected when
+fits. Furthermore, pruning outliers was attempted as well, but still yielded bad matches. The outlier critera used for
+this determination is the common interquartile range approach, where values are rejected when
 they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.
@@
 is. For instance, do cars travelling on Peachtree St. have similar travel times, despite their arrival time, or does
-their arrival drastically lengthen their time to travel the stretch of road we are modeling.
+their arrival drastically lengthen their time to travel the stretch of road we are modeling. Furthermore, we might also
+ask if the light timing has that much of an effect at all, and if it does what is the ``critical point`` at which the
+system no longer can serve incoming cars fast enough?
 
@@
 so lane changes are possible. However, we will make the assumption that drivers by this point have already made any
-discretionary lane changes and traffic is evened out between lanes. Also, we will assume left turning drivers will be
-in the left lane already. Similarly, right turning drivers will already be in the right lane.
+discretionary lane changes and traffic is evened out between lanes. This allows us to consider only a single lane at a
+time, since the second lane is identical to the first after these assumptions.
+
+We also assume here that cars are entering at the same velocity, based off of the urban speed limit of 30 mph set by
+the state of Georgia (CITE). Furthermore, we assume a single sized car, based off of the average from the NGSIM data.
+Having a constant speed and length allows us to simplify the calculation of stop points for cars entering the
+intersection. Although these simplifications help, dealing with acceleration is an even more complex matter. In this
+case it is both a mixture of both discrete (the light signal) and continuous elements (the car's velocity). By choosing
+an event-driven approach we are forced to simplify acceleration, since we cannot accurately represent a continuous
+system in this manner.
+
+Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empircal
+distribution to model the input distribution of cars. Although this is an "easy" choice, it comes with the assumption
+that both the measured data is representative of the true input distribution of cars, and that there is no measurement
+error in the data collected.
 

### 2015-02-27 project1/checkpoint/main.tex
Adding in acceleration model

@@
 road, along with a model for accelerating from a stop.
+
+\subsubsection{Acceleration Model}
+Bonneson \cite{bonneson} noted a property of car platoon acceleration: Acceleration decreases linearly with velocity.
+The formulation he provides is a differential equation:
+\[
+\frac{dV}{dt} = A_{m} \left( 1 - \frac{V}{V_{m}} \right)
+\]
+Where $A_m$ and $V_m$ where the empirically calculated maximum acceleration and velocity of cars. The acceleration was
+determined to be roughly 6.63 $\frac{ft}{s^2}$ and the maximum velocity simply the speed limit in the area (30 mph).
+This equation is indeed solvable by hand and gives us a closed form for the position of car (assuming an initial velocity of 0):
+\[
+    p(t) = V_{m}t - \frac{V_{m}^2}{A_{m}} \left( 1 - e^{-\frac{A_{m}t}{V_{m}} } \right)
+\]
+
+Assuming all cars have length $L$, the $n$th car in the queue will have to travel $L(n-1)$ feet in order to cross the
+intersection. Given our closed form above, in order to know how long it will be before a car passes we'll need to solve
+\[
+p(t) = L(n-1)
+\]
+Unfortunately, this must be done numerically. The function is fairly well behaved, and will in general have two roots
+(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method will converge
+within only a few steps. (SHOW GRAPH)
+
+Given this formulation, we can use this method to calculate when a given car will pass through the intersection,
+assuming it was at a complete stop. We'll use this information when enqueueing the "exit" event for a car.
 

### 2015-02-27 project1/checkpoint/main.tex
Cleaning a little

@@
 
-Given the difficulties with modelling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
-and simply plan to use the observed frequency distribution as our input distribution. As the project progresses we can
-refine this model.
+Given the difficulties with modeling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
+and simply plan to use the observed frequency distribution as our input distribution.
 

### 2015-02-27 project1/checkpoint/main.tex
Adding RNG discussion

@@
 
+Alongside the NGSIM data set \cite{ngsim}, we are also dependent on the browser for much of our runtime configuration.
+In particular, we rely on the random number generation capabilities of the browser for our entire source of randomness.
+JavaScript provides a built-in implementation \cite{ecmascriptmath} \texttt{Math.random()}, which provides a uniformly
+distributed random number between 0 and 1. The standard also specifies that the implementation is vendor-dependent.
+This means that the "randomness" of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
+a web page to calculate a simple $\chi^2$ statistic for a sample \cite{rngtester}. I've tested on Safari, Firefox, and
+Chrome with generally solid results, despite the fact that their implementations are different
+\cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
+with the native browser implementation of random numbers. However, these issues are not terribly important for a
+simulation of this simplicity.
+
 \subsection{Simulation Output}

### 2015-02-27 project1/checkpoint/main.tex
Adding in the acceleration graph
Citing the GA Code

@@
 Unfortunately, this must be done numerically. The function is fairly well behaved, and will in general have two roots
-(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method will converge
-within only a few steps. (SHOW GRAPH)
+(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method
+will converge within only a few steps. \ref{fig:acceleration} demonstrates some properties of this function
+$0 = p(t) - L(n-1)$, one being
+that as we move further back in the queue, the timing between cars approaches a constant. This is a result of cars
+have more space to accerate to full speed.
+
+\begin{figure}
+\begin{center}  
+\includegraphics[width=4in]{../acceleration.pdf}  
+\caption{\small \sl $p(t) - L(n-1)$ for $n \in \{0, 1, 2, 3, 4\}$ \label{fig:acceleration}}  
+\end{center}  
+\end{figure}
 
@@
 We also assume here that cars are entering at the same velocity, based off of the urban speed limit of 30 mph set by
-the state of Georgia (CITE). Furthermore, we assume a single sized car, based off of the average from the NGSIM data.
+the state of Georgia \cite{gacode}. Furthermore, we assume a single sized car, based off of the average from the NGSIM data.
 Having a constant speed and length allows us to simplify the calculation of stop points for cars entering the

### 2015-02-27 project1/checkpoint/main.tex
Adding average car length
Adding a summary section a few citations

@@
 consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
-have the same length. These values can be calculated from the NGSIM data set \cite{ngsim}.
+have the same length. These values can be calculated from the NGSIM data set \cite{ngsim} (which yields an average car length of 16.407).
 
@@
 \cite{browserify} was used to bundle the code into a single file. Minification was done with Uglify.js \cite{uglify}.
+Graphing is performed via inline SVG \cite{svg}.
 
 \section{Current Build and Running Instructions}
-The simulation is currently deployed at \cite{aprilandchip} and currently runs a fixed number of cars through stepping
-at 1 second real time being equal to 100ms. Eventually, this will be adjustable in real-time, and better visualized.
-Cars are generated regularly at 10 (+ an uniformly random number between 0 and 1) second intervals. Light timings are
-currently hard-coded as well. Currently, we're only modeling a single direction, and ignoring turns, and variations
-between cars.
+The simulation is currently deployed at \cite{aprilandchip} and provides a UI that allows the light change intervals to
+be updated in real time. Moreover, the simulation speed can be adjusted in real-time as well. Worth noting in the UI is
+a log of all events that are passing through the system. This is live filterable. Also, a visualization of the current
+signal state, along with the current wait times for the first 5 cars in line is provided. Statistics are provided in
+real-time, providing both the observed input distribution of car inter-arrival times, and the output distribution of
+car wait times.
 
@@
 \texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
-simulation.
+simulation. To rerun the simulation, simply refresh the browser
+
+\section{Results}
+Since light timing is adjustable in real-time, the actual average wait times can vary (even disregarding randomness).
+However, a few general trends tend to appear. First off, the distribution of wait times is bimodal. This can be
+somewhat observed in the NGSIM data set. This appears to be a result of the dichotomy of cars arriving when the light
+is green, versus when it is red. Cars that arrive during a green light in general do not have to wait long at all
+before they can cross the intersection. However, cars that arrive during a red light will be forced to wait, often
+behind other cars. This causes their wait time to be much higher since they actually had to queue up before they could
+exit the system. Unsurprisingly, most cars that catch the red light end up waiting around the duration of red time.
+
+A second observation is the system saturation. There is a point where the light timing is so terrible that the system
+cannot push cars through as fast as they come. In queueing theory terms, the system is referred to as being
+``oversaturated``. After tweaking some values, this can be triggered by setting the red duration to 60 seconds and
+slowing adjusting the yellow and green down (my testing resulted in oversaturation around yellow and green with 1 and
+25 second durations respectively).
+
+\section{Conclusion}
 

### 2015-02-27 project1/checkpoint/main.tex
Adding a conclusion and summary

@@
 \section{Conclusion}
+In summary, I've built a web-based simulation of the 12th and Peachtree intersection in both the North and South
+directions. The NGSIM dataset was used for gathering estimations about input rates, and various general statistics about
+the cars on Peachtree. Some simplifications were made in order to accomodate lack of data, domain knowledge, and time
+(among other things). Furthermore, some of the design choices have their sets of caveats which were noted when building
+the simulation. However, the simulation is real-time and does yield relatively intuitive results.
 

### 2015-02-28 project1/checkpoint/main.tex
Final tidying

@@
 intersection's timings are time-based, rather than sensor-based, which can be seen in the data set provided with the
-project description and is also available online \cite{ngsim}. We assume that the current light timing provided here
+project description and is also available online \cite{ngsim}. We assume that the current light timings provided here
 are optimal, but because we'll have a computer simulation in the end we have free reign in experimenting with these
@@
 \]
-Where $A_m$ and $V_m$ where the empirically calculated maximum acceleration and velocity of cars. The acceleration was
+where $A_m$ and $V_m$ are the empirically calculated maximum acceleration and velocity of cars. The acceleration was
 determined to be roughly 6.63 $\frac{ft}{s^2}$ and the maximum velocity simply the speed limit in the area (30 mph).
@@
 \]
-Unfortunately, this must be done numerically. The function is fairly well behaved, and will in general have two roots
+Unfortunately, this has no closed form and must be done numerically. The function is fairly well behaved, and will in general have two roots
 (only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method
-will converge within only a few steps. \ref{fig:acceleration} demonstrates some properties of this function
+will converge within only a few steps. Figure \ref{fig:acceleration} demonstrates some properties of this function
 $0 = p(t) - L(n-1)$, one being
 that as we move further back in the queue, the timing between cars approaches a constant. This is a result of cars
-have more space to accerate to full speed.
+have more space to accelerate to full speed.
 
@@
 
-The car model will listen for this event as well. When a currently traveling car sees a red light, it will calculate
-\emph{where} it needs to stop, and \emph{when} it will come to a stop. This will be done using a deceleration model
-similar to \cite{deceleration}. If the light turns to green, the car instead needs to start accelerating (using one of
-the models mentioned above). The model will adapt to light changes during movement by recalculating its estimated
-exit time of the system.
+Furthermore, the queue at the intersection will need to listen for this event as well. In essense, when the light
+changes, the intersection will need to switch between holding cars and actually passing them through. If the light has
+turned to green, it will need to one-by-one discharge cars based on the above acceleration model. If the light is red,
+the queue will need to appropriate insert the car to await the car's turn to pass through the intersection.
 
 \subsection{Simulation Input}
-The two main sources of input into this simulation are the cars entering out section of Peachtree St, and the light
+The two main sources of input into this simulation are the cars entering our section of Peachtree St, and the light
 timing of the system. The combination of these two variables are what influence the smoothness
@@
 Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
-fits. Furthermore, pruning outliers was attempted as well, but still yielded bad matches. The outlier critera used for
+fits (verified by $\chi^2$ goodness-of-fit tests. Furthermore, pruning outliers was attempted as well, but still
+yielded bad matches. The outlier critera used for
 this determination is the common interquartile range approach, where values are rejected when
@@
 
-Given the difficulties with modeling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
+Given the difficulties with modeling we'll take a simpler approach and assume the true distribution is modeled by what was observed,
 and simply plan to use the observed frequency distribution as our input distribution.
@@
 authored the \texttt{chi-squared-test} module \cite{chiSquaredNpm}, and the \texttt{gamma-distribution}
-\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used a build tool for automating tasks (including building this paper).
+\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used as a build tool for automating tasks (including building this paper).
 To run some of the data analysis code I wrote, you'll need to install and run gulp. Instructions for doing so are
@@
 JavaScript provides a built-in implementation \cite{ecmascriptmath} \texttt{Math.random()}, which provides a uniformly
-distributed random number between 0 and 1. The standard also specifies that the implementation is vendor-dependent.
-This means that the "randomness" of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
-a web page to calculate a simple $\chi^2$ statistic for a sample \cite{rngtester}. I've tested on Safari, Firefox, and
-Chrome with generally solid results, despite the fact that their implementations are different
-\cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
+distributed random number between 0 and 1. The standard specifies that the implementation is vendor-dependent.
+This means that the ``randomness'' of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
+a web page to calculate a simple $\chi^2$ statistic for a sample of \texttt{Math.random()} \cite{rngtester}. I've tested
+\texttt{Math.random()} on Safari, Firefox, and Chrome with generally solid results, despite the fact that their implementations
+are different \cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
 with the native browser implementation of random numbers. However, these issues are not terribly important for a
@@
 
-Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empircal
-distribution to model the input distribution of cars. Although this is an "easy" choice, it comes with the assumption
+Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empirical
+distribution to model the input distribution of cars. Although this is an ``easy'' choice, it comes with the assumption
 that both the measured data is representative of the true input distribution of cars, and that there is no measurement
@@
 Since we want to gather metrics regarding the throughput of cars, we'll also provide a listener class
-\texttt{ThroughputStatistics} that will listen for cars to complete their journey through the system. This will
+\texttt{StatCollection} that will listen for cars to complete their journey through the system. This will
 be reported in the UI in real time.
@@
 \texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
-simulation. To rerun the simulation, simply refresh the browser
+simulation. To rerun the simulation, simply refresh the page.
 

### 2015-03-14 project2/wmata/pingApi.sh
Adding a single script for hitting the WMATA API

@@ -1,0 +1,16 @@
+#!/bin/bash
+
+output_dir="$1"
+
+if [ -z "$output_dir" ]; then
+    output_dir="."
+fi
+
+# format the current date+time
+output_file=`date +"%s.json"`
+
+# pluck my subscription key, stripping newlines and things
+subscription_key=`cat WMATA.txt`
+
+# Hit the api, using my subscription key, and get ALL stations.
+curl -X GET "https://api.wmata.com/StationPrediction.svc/json/GetPrediction/All?api_key=$subscription_key" > $output_dir/$output_file

### 2015-03-14 project2/wmata/pingApi.sh
Adding nice args and stuff

@@ -1,16 +1,27 @@
 #!/bin/bash
 
-output_dir="$1"
+# cron schedule (all the time): * * * * * *
 
-if [ -z "$output_dir" ]; then
-    output_dir="."
-fi
+output_dir="."
+subscription_key="NOPE"
+
+while getopts "h?o:?k:" opt; do
+    case "$opt" in
+    h|\?)
+        echo "Usage: ./pingApi.sh -k <API_KEY> [-o output_directory]"
+        exit 0
+        ;;
+    o)
+        output_dir=$OPTARG
+        ;;
+    k)
+        subscription_key="$OPTARG"
+        ;;
+    esac
+done
 
 # format the current date+time
 output_file=`date +"%s.json"`
-
-# pluck my subscription key, stripping newlines and things
-subscription_key=`cat WMATA.txt`
 
 # Hit the api, using my subscription key, and get ALL stations.
 curl -X GET "https://api.wmata.com/StationPrediction.svc/json/GetPrediction/All?api_key=$subscription_key" > $output_dir/$output_file

### 2015-03-14 project2/wmata/insertIntoMongo.sh
First stab at a mongo insert scrip

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+fileToInsert="$1"
+
+if [ -z "$fileToInsert" ]; then
+    echo "Missing a file!"
+    exit 1
+fi
+
+# convert the json by appending the timestamp of the file on first
+timestamp=`echo $fileToInsert | cut -d'.' -f1`
+document=`cat $fileToInsert | sed "s/^{/{\"timestamp\":$timestamp,/"`
+
+# now append into mongo
+echo "use metro; db.trains.insert($document);" | mongo

### 2015-03-14 project2/wmata/insertIntoMongo.sh
Using mongo import now

@@ -9,7 +9,10 @@ fi
 
 # convert the json by appending the timestamp of the file on first
 timestamp=`echo $fileToInsert | cut -d'.' -f1`
-document=`cat $fileToInsert | sed "s/^{/{\"timestamp\":$timestamp,/"`
+cat $fileToInsert | sed "s/^{/{\"timestamp\":$timestamp,/" > document.json
 
 # now append into mongo
-echo "use metro; db.trains.insert($document);" | mongo
+mongoimport --db metro --collection trains < document.json
+
+# cleanup
+rm document.json

### 2015-03-14 project2/wmata/pingApi.sh
Making sure the output directory is there

@@ -23,5 +23,8 @@ done
 # format the current date+time
 output_file=`date +"%s.json"`
 
+# make sure the output directory is there
+mkdir -p $output_dir
+
 # Hit the api, using my subscription key, and get ALL stations.
 curl -X GET "https://api.wmata.com/StationPrediction.svc/json/GetPrediction/All?api_key=$subscription_key" > $output_dir/$output_file

### 2015-03-14 project2/wmata/appendTimestamp.py
Adding a helper script for appending timestamps

@@ -1,0 +1,20 @@
+import sys, json
+
+def appendTimestampToSingleTrain(train, timestamp):
+    train['timestamp'] = timestamp
+    return train
+
+if len(sys.argv) is 1:
+    sys.exit(1)
+
+timestamp = int(sys.argv[1])
+
+inputJson = sys.stdin.read().replace('\n', '')
+
+inputJson = json.loads(inputJson)
+
+trains = []
+for train in inputJson['Trains']:
+    trains.append(appendTimestampToSingleTrain(train, timestamp))
+
+print(json.dumps(trains))

### 2015-03-14 project2/wmata/insertIntoMongo.sh
Using the helper script to do the appending

@@ -9,10 +9,10 @@ fi
 
 # convert the json by appending the timestamp of the file on first
 timestamp=`echo $fileToInsert | cut -d'.' -f1`
-cat $fileToInsert | sed "s/^{/{\"timestamp\":$timestamp,/" > document.json
+cat $fileToInsert | python appendTimestamp.py $timestamp > documents.json
 
 # now append into mongo
-mongoimport --db metro --collection trains < document.json
+mongoimport --db metro --collection trains < documents.json
 
 # cleanup
 rm document.json

### 2015-03-14 project2/wmata/insertIntoMongo.sh
Fixing the filename

@@ -15,4 +15,4 @@ cat $fileToInsert | python appendTimestamp.py $timestamp > documents.json
 mongoimport --db metro --collection trains < documents.json
 
 # cleanup
-rm document.json
+rm documents.json

### 2015-03-14 project2/wmata/insertIntoMongo.sh
Telling mongo that there's multiple records

@@ -12,7 +12,7 @@ timestamp=`echo $fileToInsert | cut -d'.' -f1`
 cat $fileToInsert | python appendTimestamp.py $timestamp > documents.json
 
 # now append into mongo
-mongoimport --db metro --collection trains < documents.json
+mongoimport --db metro --collection trains --jsonArray < documents.json
 
 # cleanup
 rm documents.json

### 2015-03-28 project2/queries/post_query_cleanup.sh
Don't forget to smooth

@@ -3,7 +3,7 @@
 # For train frequencies, we want interarrival times, so we do running diffs
 for file in `ls train_frequencies/*.csv`; do
     parameter_set_name=`basename $file | sed 's/.csv//g'`
-    cat $file | ../utils/train_estimate | ../utils/cluster 180 | ../utils/running_difference > ${parameter_set_name}_diffs.csv
+    cat $file | ../utils/train_estimate | ../utils/cluster 180 | ../utils/running_difference | ../utils/filter_outliers | ../utils/bin 50 | ../utils/smooth 4 > ${parameter_set_name}_diffs.csv
 done
 
 # bundle them up

### 2015-03-28 project2/queries/time_between_stations/time_between_stations.js
Making query generic

@@ -1,5 +1,17 @@
-db.trains.find({ Line: "BL", LocationCode: "C02", DestinationCode: "G05", Min: "1" }).forEach(function(localTrain) {
-    var neighborTrain = db.trains.find({ Line: "BL", LocationCode: "C01", DestinationCode: "G05", timestamp: train.timestamp })[0];
+var departingStation = {
+    Line: "${Line}",
+    LocationCode: "${DepartingStation}",
+    DestinationCode: "${DestinationCode}",
+    Min: "1"
+};
+
+var arrivingStation = {
+    Line: "${Line}",
+    LocationCode: "${ArrivingStation}",
+    DestinationCode: "${DestinationCode}"
+
+db.trains.find(departingStation).forEach(function(localTrain) {
+    arrivingStation.timestamp = localTrain.timestamp;
+    var neighborTrain = db.trains.find(arrivingStation)[0];
     print(parseInt(neighborTrain.Min) - 1);
 });
-

### 2015-03-28 project2/queries/time_between_stations/generate_queries.sh
Stubbing out the query generator

@@ -1,0 +1,81 @@
+#!/bin/bash
+
+set -e
+
+#########################################
+# Skips the first $1 lines passed to it #
+#########################################
+skip() {
+    # skip the lines, if we actually have some to skip
+    if [ "$1" -ne "0" ]; then
+        for i in $(seq 1 $1); do
+            read line
+        done
+    fi
+
+    # just echo the rest
+    while read line; do
+        echo $line
+    done
+}
+
+#######################################################################
+# Gets a trains endpoint by color $1, and direction $2 (East or West) #
+#######################################################################
+get_train_endpoint() {
+    cat ../../data/endpoints.csv | grep $1 | grep $2 | cut -d',' -f1
+}
+
+########################################
+# Gets the station for the $1-th train #
+########################################
+get_station_by_index() {
+    cat ../../data/stations.csv | skip $1 | head -n1
+}
+
+#############################################
+# Gets the station code for the $1-th train #
+#############################################
+get_station_code_by_index() {
+    get_station_by_index $1 | cut -d',' -f1
+}
+
+##############################################
+# Gets a file friendly station name by index #
+##############################################
+get_station_name_by_index() {
+    get_station_by_index $1 | cut -d',' -f2 | sed "s/ /_/g" | sed "s/\'//g" | tr "[:upper:]" "[:lower:]"
+}
+
+station_count=`cat ../../data/stations.csv | wc -l | tr -d "[:blank:]"`
+track_section_count=`expr "$station_count" - 1`
+
+# for all stations
+for station_index in $(seq 0 $(($track_section_count-1))); do
+    next_station_index=`expr "$station_index" + 1`
+    east_station=`get_station_name_by_index $station_index`
+    west_station=`get_station_name_by_index $next_station_index`
+
+    # for all colors
+    for line in BL SV OR; do
+        west_destination=`get_train_endpoint West $line`
+        # build the westbound properties file
+        westbound_file="${west_station}_west_${line}.properties"
+        echo $westbound_file
+        touch $westbound_file
+        echo "Line=$line" >> $westbound_file
+        echo "DepartingStation=$east_station" >> $westbound_file
+        echo "ArrivingStation=$west_station" >> $westbound_file
+        echo "DestinationCode=$west_destination" >> $westbound_file
+        
+        east_destination=`get_train_endpoint East $line`
+        # build the westbound properties file
+        eastbound_file="${west_station}_east_${line}.properties"
[…]

### 2015-03-28 project2/queries/time_between_stations/generate_queries.sh
Fixing some query generation

@@ -50,32 +50,36 @@ get_station_name_by_index() {
 station_count=`cat ../../data/stations.csv | wc -l | tr -d "[:blank:]"`
 track_section_count=`expr "$station_count" - 1`
 
+rm -f *.properties
+
 # for all stations
 for station_index in $(seq 0 $(($track_section_count-1))); do
     next_station_index=`expr "$station_index" + 1`
-    east_station=`get_station_name_by_index $station_index`
-    west_station=`get_station_name_by_index $next_station_index`
+    east_station_name=`get_station_name_by_index $station_index`
+    east_station_code=`get_station_code_by_index $station_index`
+    west_station_name=`get_station_name_by_index $next_station_index`
+    west_station_code=`get_station_code_by_index $next_station_index`
 
     # for all colors
     for line in BL SV OR; do
         west_destination=`get_train_endpoint West $line`
         # build the westbound properties file
-        westbound_file="${west_station}_west_${line}.properties"
+        westbound_file="${west_station_name}_west_${line}.properties"
         echo $westbound_file
         touch $westbound_file
         echo "Line=$line" >> $westbound_file
-        echo "DepartingStation=$east_station" >> $westbound_file
-        echo "ArrivingStation=$west_station" >> $westbound_file
+        echo "DepartingStation=$east_station_code" >> $westbound_file
+        echo "ArrivingStation=$west_station_code" >> $westbound_file
         echo "DestinationCode=$west_destination" >> $westbound_file
         
         east_destination=`get_train_endpoint East $line`
         # build the westbound properties file
-        eastbound_file="${west_station}_east_${line}.properties"
+        eastbound_file="${east_station_name}_east_${line}.properties"
         echo $eastbound_file
         touch $eastbound_file
         echo "Line=$line" >> $eastbound_file
-        echo "DepartingStation=$west_station" >> $eastbound_file
-        echo "ArrivingStation=$east_station" >> $eastbound_file
+        echo "DepartingStation=$west_station_code" >> $eastbound_file
+        echo "ArrivingStation=$east_station_code" >> $eastbound_file
         echo "DestinationCode=$east_destination" >> $eastbound_file
     done
 done

### 2015-03-29 project2/queries/time_between_stations/time_between_stations.js
Typo in the query.

@@ -9,6 +9,7 @@ var arrivingStation = {
     Line: "${Line}",
     LocationCode: "${ArrivingStation}",
     DestinationCode: "${DestinationCode}"
+};
 
 db.trains.find(departingStation).forEach(function(localTrain) {
     arrivingStation.timestamp = localTrain.timestamp;

### 2015-03-30 project2/sim/gulpfile.js
Stubbing out some of the initial code

@@ -1,0 +1,21 @@
+var gulp = require('gulp');
+var del = require('del');
+var coffeeify = require('gulp-coffeeify');
+var uglify = require('gulp-uglify');
+
+gulp.task('bundle', function() {
+    return gulp.src('coffee/main.coffee')
+        .pipe(coffeeify())
+        .pipe(uglify({ "source-map" : "main.js.map" }))
+        .pipe(gulp.dest('.'));
+});
+
+gulp.task('clean', function(cb) {
+    del('main.js', cb);
+});
+
+gulp.task('default', ['bundle']);
+
+gulp.task('watch', ['bundle'], function() {
+    gulp.watch('coffee/*', 'bundle');
+});

### 2015-03-30 project2/paper/main.tex
Stubbing out some latex stuff I'll need

--- 2015-02-27 project1/checkpoint/main.tex
@@
 
-\title{Project 1 Checkpoint}
+\title{Project 2}
 \author{Chip Bell}
-\date{February 6th, 2015}
+\date{March 31, 2015}
 \maketitle
-
-\section{Problem Description}
-Traffic simulation is becoming increasing important due to the expansion and growth of cities. Long-standing
-infrastructure is forced to adapt in order to accomodate greater amounts of traffic, despite the fact that in many
-cases the roads cannot be modified. This can be a result of legislation, existing buildings, and private property among
-other things. So as a result, cities must instead optimize the variables they can control. In this case, we consider
-signal light timing for the 12th St. and Peachtree Street intersection. In particular,
-we explore the effects that light timing has on traffic flow, given a source input distribution of cars.
-
-\section{The Conceptual Model}
-In order to create a representative software simulation of the aforementioned system, we first begin by designing a
-conceptual version of this model from which we can plan our implementation. We'll borrow heavily from the results and data
-in \cite{cts12} when formulating our model.
-Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
-we'll utilize the publisher-subscriber pattern (see \cite{pubsub}) to promote better decoupling.
-Given this architecture, our first step will be to identify the entities present in our model, and then consider their
-interactions (or the inherent ``activies'') within the simulation. This will allow us then to design the events that
-are important and relevant to the simulation.
-
-\subsection{Entities}
-The first entity, unsurprisingly, is a single car. Our representation of a car is clearly less detailed than the real
-thing, but rather focuses on variables that influence it's travel time from its starting location to destination, along
-with the variables that influence neighboring cars. For our simulation, we'll track the position and destination of the
-cars. Probably the most primary value associated with a car is its arrival time, since that dictates when and where it
-will fall in the queue of cars at the intersection.
-
-Another entity to consider is the intersection itself. It's changing in time constantly, and those changes affect
-every entity within the simulation. The most important parameters to consider here are the light timings. This
-intersection's timings are time-based, rather than sensor-based, which can be seen in the data set provided with the
-project description and is also available online \cite{ngsim}. We assume that the current light timing provided here
-are optimal, but because we'll have a computer simulation in the end we have free reign in experimenting with these
-values and observing the outcome. Perhaps the timings are \emph{not} optimal. Perhaps, our model grossly oversimplifies
-some crucial interaction within the system. This is perhaps the most exciting part of the project.
-
-\subsection{Basic Interactions}
-From the intersection entity, we can glean an important interaction that occurs: The changing of light signal. When a
-signal changes, each car in-queue will be forced to change it's state, whether it's waiting or currently in motion.
-From an event-driven perspective, this means that any listener would need to be able to know \emph{how} the light
-changed in order to act accordingly.
-
-An intersection serves as holding point for cars traveling along the road so if a light is red, cars are forced to
-queue up at the intersection until green. This lends itself to a queue-style data structure that encorporates both
-position and acceleration. Car acceleration is a complex topic, but models have been constructed \cite{bonneson}
-\cite{herman_et_al}. These models are generally continuous, and can generally be represented as a differential equation
-due to the relationships between acceleration and velocity, as can be seen in \cite{briggs} \cite{deceleration} as well.
-Instead, we will simply consider the various events that trigger changes in behavior. For instance, a real driver is
-constantly observing their environment (hopefully) and making minute adjustments the entire time. However, we'll take
-a greatly simplified approach and simply precalculate an estimate of how long a car takes to traverse the section of
-road, along with a model for accelerating from a stop.
-
-\subsubsection{Acceleration Model}
-Bonneson \cite{bonneson} noted a property of car platoon acceleration: Acceleration decreases linearly with velocity.
-The formulation he provides is a differential equation:
-\[
-\frac{dV}{dt} = A_{m} \left( 1 - \frac{V}{V_{m}} \right)
-\]
-Where $A_m$ and $V_m$ where the empirically calculated maximum acceleration and velocity of cars. The acceleration was
-determined to be roughly 6.63 $\frac{ft}{s^2}$ and the maximum velocity simply the speed limit in the area (30 mph).
-This equation is indeed solvable by hand and gives us a closed form for the position of car (assuming an initial velocity of 0):
-\[
-    p(t) = V_{m}t - \frac{V_{m}^2}{A_{m}} \left( 1 - e^{-\frac{A_{m}t}{V_{m}} } \right)
-\]
-
-Assuming all cars have length $L$, the $n$th car in the queue will have to travel $L(n-1)$ feet in order to cross the
-intersection. Given our closed form above, in order to know how long it will be before a car passes we'll need to solve
-\[
-p(t) = L(n-1)
-\]
-Unfortunately, this must be done numerically. The function is fairly well behaved, and will in general have two roots
-(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method
-will converge within only a few steps. \ref{fig:acceleration} demonstrates some properties of this function
-$0 = p(t) - L(n-1)$, one being
-that as we move further back in the queue, the timing between cars approaches a constant. This is a result of cars
-have more space to accerate to full speed.
-
-\begin{figure}
-\begin{center}  
-\includegraphics[width=4in]{../acceleration.pdf}  
-\caption{\small \sl $p(t) - L(n-1)$ for $n \in \{0, 1, 2, 3, 4\}$ \label{fig:acceleration}}  
-\end{center}  
-\end{figure}
-
-Given this formulation, we can use this method to calculate when a given car will pass through the intersection,
-assuming it was at a complete stop. We'll use this information when enqueueing the "exit" event for a car.
-
-\subsection{Events}
-Based on our observations from above, we can now construct a set of events that are pertinent to this application.
-
-First off, signal changes consititute the major event in the system. When a signal changes, the event must contain
-\emph{when} the signal changed, and \emph{what} the signal changed to. The light itself will listen for this event
-and will enqueue the next signal change based on the signal change time. These signal changes could be enqueued
-completely upfront, but allowing this to be done on the fly allows the user to tweak these values during the simulation.
-
-The car model will listen for this event as well. When a currently traveling car sees a red light, it will calculate
-\emph{where} it needs to stop, and \emph{when} it will come to a stop. This will be done using a deceleration model
-similar to \cite{deceleration}. If the light turns to green, the car instead needs to start accelerating (using one of
-the models mentioned above). The model will adapt to light changes during movement by recalculating its estimated
-exit time of the system.
-
-\subsection{Simulation Input}
-The two main sources of input into this simulation are the cars entering out section of Peachtree St, and the light
-timing of the system. The combination of these two variables are what influence the smoothness
-of traffic flow to the actual drivers in the system.
-
-We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
-consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
-have the same length. These values can be calculated from the NGSIM data set \cite{ngsim} (which yields an average car length of 16.407).
-
-\begin{figure}
-\begin{center}  
-\includegraphics[width=4in]{../northbound.png}  
-\caption{\small \sl Northbound Arrival Intervals by Order of Occurrence\label{fig:northbound_arrival}}  
-\end{center}  
-\end{figure}
-
-\begin{figure}  
-\begin{center}
-\includegraphics[width=4in]{../southbound.png}
-\caption{\small \sl Southbound Arrival Intervals by Order of Occurrence\label{fig:southbound_arrival}}  
-\end{center}  
-\end{figure}
-
-The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
-northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
-Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
-fits. Furthermore, pruning outliers was attempted as well, but still yielded bad matches. The outlier critera used for
-this determination is the common interquartile range approach, where values are rejected when
-they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.
-
-Given the difficulties with modeling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
-and simply plan to use the observed frequency distribution as our input distribution.
-
-The current light timing of the system is known, and provided in \cite{ngsim}, but we'll allow the user to set
-those values dynamically via a UI.
-
-\subsubsection{Data Analysis}
-In the \texttt{data\_processing} folder, most of code used for distribution fitting and statistics can be found there.
-The data analysis code was written in JavaScript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
-to manage dependencies and those dependencies can be seen in the \texttt{package.json} file. Of those packages, I have
-authored the \texttt{chi-squared-test} module \cite{chiSquaredNpm}, and the \texttt{gamma-distribution}
-\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used a build tool for automating tasks (including building this paper).
-To run some of the data analysis code I wrote, you'll need to install and run gulp. Instructions for doing so are
-provided in the final section of this paper.
-
-Alongside the NGSIM data set \cite{ngsim}, we are also dependent on the browser for much of our runtime configuration.
-In particular, we rely on the random number generation capabilities of the browser for our entire source of randomness.
-JavaScript provides a built-in implementation \cite{ecmascriptmath} \texttt{Math.random()}, which provides a uniformly
-distributed random number between 0 and 1. The standard also specifies that the implementation is vendor-dependent.
-This means that the "randomness" of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
-a web page to calculate a simple $\chi^2$ statistic for a sample \cite{rngtester}. I've tested on Safari, Firefox, and
-Chrome with generally solid results, despite the fact that their implementations are different
-\cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
-with the native browser implementation of random numbers. However, these issues are not terribly important for a
-simulation of this simplicity.
-
-\subsection{Simulation Output}
-The most important statistic we'll want to collect during the simulation is the thoughput of the simulation. We'll
-want to observe how the light timing affects the average travel time for cars. It will also to see how variable this
-is. For instance, do cars travelling on Peachtree St. have similar travel times, despite their arrival time, or does
-their arrival drastically lengthen their time to travel the stretch of road we are modeling. Furthermore, we might also
-ask if the light timing has that much of an effect at all, and if it does what is the ``critical point`` at which the
-system no longer can serve incoming cars fast enough?
-
-\subsection{Simplifications and Assumptions}
-Probably the most obvious "flaw" in such a design are the simplifications made. Details can certainly be left out
-purposely for many reasons such as time constraints, lack of supporting data to model a extra feature, or simply
-unneccesary complexity.
-
-In this case, certainly the most prominent would be the ``isolation'' of the model. We are only considering an
-idealized random input distribution of cars from both North and South of the 12th St. intersection. However, this is
-certainly not the case, since this input distribution is affected by countless other factors like the surrounding
-lights, cars passing along 12th at that intersection, and even simply time of day. We assume in this simulation is
-perpetually between 4:00PM to 4:15PM in Atlanta (a frightening thought indeed), simply due to the fact that we only
-have data for that time range. All in all, we assume the conditions were fairly identical to that in which the 
-aforementioned CTS study was made.
-
-Another simplification to consider is lane changing. Lane changing models do exist \cite{lanechanging} \cite{lanechanging2},
-but can be complex, requiring modeling of driver choices and gap acceptance for drivers. Peachtree is a two lane road,
-so lane changes are possible. However, we will make the assumption that drivers by this point have already made any
-discretionary lane changes and traffic is evened out between lanes. This allows us to consider only a single lane at a
-time, since the second lane is identical to the first after these assumptions.
-
-We also assume here that cars are entering at the same velocity, based off of the urban speed limit of 30 mph set by
-the state of Georgia \cite{gacode}. Furthermore, we assume a single sized car, based off of the average from the NGSIM data.
-Having a constant speed and length allows us to simplify the calculation of stop points for cars entering the
-intersection. Although these simplifications help, dealing with acceleration is an even more complex matter. In this
-case it is both a mixture of both discrete (the light signal) and continuous elements (the car's velocity). By choosing
-an event-driven approach we are forced to simplify acceleration, since we cannot accurately represent a continuous
-system in this manner.
-
-Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empircal
-distribution to model the input distribution of cars. Although this is an "easy" choice, it comes with the assumption
-that both the measured data is representative of the true input distribution of cars, and that there is no measurement
-error in the data collected.
-
-\section{Simulation Software Architecture}
-Because of the event-driven architecture, the system can be though of as a set of event emitters and event listeners.
-In some cases, an object may both emit and listen for events.
-
-The first event emitter to consider is the \texttt{CarEmitter}. The \texttt{CarEmitter} uses the measured input
-distribution we have for cars and generates a set of cars randomly. It then emits these cars as individual
-\texttt{car:arrived} event. There will generally be two \texttt{CarEmitter}s, one for the North and South direction.
-
-The second emitter will be the \texttt{LightSignal}. It will emit \texttt{light:changed} events, passing the light color.
-Also, instead of enqueueing these events upfront, it will also listen for \texttt{light:changed} so that its
-changing duration can modified on the fly (using the UI).
-
-The \texttt{IntersectionQueue} holds a list of cars lined up at the intersection. It also maintains the
-current signal state, so that it can send cars directly through the intersection if the light is green. It responds
-to pretty much every event mentioned above. When the light changes, it will need update it's current signal state.
-This may mean sending some cars off of the queue and stop holding cars, or to no longer let cars through and make them
-wait. Furthermore, when a car arrives it will handle that event by deciding whether the car can pass through or needs
-to wait on the light. Also, this class emits events. Whenever a car finally crosses the intersection, it emits
-a \texttt{car:exited} event.
-
-Since we want to gather metrics regarding the throughput of cars, we'll also provide a listener class
-\texttt{ThroughputStatistics} that will listen for cars to complete their journey through the system. This will
-be reported in the UI in real time.
-
-The simulation is written in CoffeeScript \cite{coffeescript}, which is a small language that compiles to JavaScript.
-CoffeeScript removes some of the boilerplate associated with setting up prototype chains and other object-oriented
-concepts in JavaScript.
-Backbone \cite{backbone} along with Underscore \cite{underscore} and jQuery \cite{jquery} were used to provide some
-application structure, like basic eventing, and the sorted
-collections that are used to form the basis for the event queue and \texttt{IntersectionQueue}. Lastly, Browserify
-\cite{browserify} was used to bundle the code into a single file. Minification was done with Uglify.js \cite{uglify}.
-Graphing is performed via inline SVG \cite{svg}.
-
-\section{Current Build and Running Instructions}
-The simulation is currently deployed at \cite{aprilandchip} and provides a UI that allows the light change intervals to
-be updated in real time. Moreover, the simulation speed can be adjusted in real-time as well. Worth noting in the UI is
-a log of all events that are passing through the system. This is live filterable. Also, a visualization of the current
-signal state, along with the current wait times for the first 5 cars in line is provided. Statistics are provided in
-real-time, providing both the observed input distribution of car inter-arrival times, and the output distribution of
-car wait times.
-
-To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
-with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
-\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
-\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
-\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
-simulation. To rerun the simulation, simply refresh the browser
-
-\section{Results}
-Since light timing is adjustable in real-time, the actual average wait times can vary (even disregarding randomness).
-However, a few general trends tend to appear. First off, the distribution of wait times is bimodal. This can be
-somewhat observed in the NGSIM data set. This appears to be a result of the dichotomy of cars arriving when the light
-is green, versus when it is red. Cars that arrive during a green light in general do not have to wait long at all
-before they can cross the intersection. However, cars that arrive during a red light will be forced to wait, often
-behind other cars. This causes their wait time to be much higher since they actually had to queue up before they could
-exit the system. Unsurprisingly, most cars that catch the red light end up waiting around the duration of red time.
-
-A second observation is the system saturation. There is a point where the light timing is so terrible that the system
-cannot push cars through as fast as they come. In queueing theory terms, the system is referred to as being
-``oversaturated``. After tweaking some values, this can be triggered by setting the red duration to 60 seconds and
-slowing adjusting the yellow and green down (my testing resulted in oversaturation around yellow and green with 1 and
-25 second durations respectively).
-
-\section{Conclusion}
 

### 2015-03-30 project2/paper/gulpfile.js

@@ -1,0 +1,28 @@
+var gulp = require('gulp');
+var async = require('async');
+var exec = require('child_process').exec;
+var del = require('del');
+
+// Converts a string command into an async exec call, to be chained with async.series
+var asyncCommand = function(commandText) {
+    return function(callback) {
+        exec(commandText, callback);
+    };
+};
+
+gulp.task('clean', function(cb) {
+    del(['*.{aux,bbl,blg,dvi,log,pdf}'], cb);
+});
+
+
+gulp.task('default', ['clean'], function(cb) {
+    var latex = 'pdflatex -halt-on-error main';
+    var bibtex = 'bibtex main';
+    
+    async.series([
+        asyncCommand(latex),
+        asyncCommand(bibtex),
+        asyncCommand(latex),
+        asyncCommand(latex)
+    ], cb);
+});

### 2015-03-30 project2/paper/gulpfile.js
Forgot a file

@@ -11,7 +11,7 @@ var asyncCommand = function(commandText) {
 };
 
 gulp.task('clean', function(cb) {
-    del(['*.{aux,bbl,blg,dvi,log,pdf}'], cb);
+    del(['*.{aux,bbl,blg,dvi,log,pdf,out}'], cb);
 });
 
 

### 2015-03-30 project2/paper/main.tex
Getting builds working, and adding a few citations I'll probably need
Adding summary section

@@
 
+\section{Problem Description}
+The Washington Metropolitan Area Transit Authority's Metrorail system in many respects is a driving factor in the
+expansion of the city. With Washington's fixed size, reliable and cost-efficient transportation for metro area
+residents is crucial for maintaining the city's economy. However, this is a difficult task in that there are many
+variables involved, such as human error, mechanical failure, and even acts of nature.
+
+Therefore for this project, I will be simulating the metrorail system using data collected from the WMATA API
+\cite{wmataapi}, which provides realtime estimates of train arrival. Furthermore, I will simulate the impact that track
+closures have on train throughput.
+
 \bibliographystyle{plain}

### 2015-03-31 project2/paper/main.tex
Fleshing out a little

@@
 
+\section{Previous Work}
+Many large cities have rapid transit systems, so considerable research time has been invested on building reliable
+subways. Commercial software such as OpenTrack \cite{opentrack} have been developed, and even video games such as Train
+Simulator \cite{trainsimulator} and Open Rails \cite{openrails}. Algorithms for handling single, double and tracks exist, such as the basic intuitive
+algorithm for sharing tracks between trains presented by Fiorini and Botter \cite{fioroni}. Dessouky and Leachman
+\cite{dessouky_leachman_95} provide an event-driven algorithm for single and double track rail networks, incorporating
+acceleration. Lu et al \cite{quan_lu} expand the model to triple tracks while enhancing the acceleration model to
+handle multiple track speed limits.
+
+\section{Basic Architecture}
+
+\section{Data Collection}
+
+\section{Caveats and Limitations}
+
+\section{Technology Used}
+
+\section{Remaining Work}
+
 \bibliographystyle{plain}

### 2015-03-31 project2/paper/main.tex
Adding some figures and data analysis stuff

@@
 \section{Basic Architecture}
+The nature of trains lends itself to a model that incorporates resource sharing. For instance, only a single train can
+occupy a particular section of track at a given moment. Furthermore, if two trains are sharing a track, they cannot
+be going in different directions unless they have already past each other, lest they collide. This issue is addressed
+by Fioroni \cite{fioroni}.
+
+A common occurrence in the metro system is that between two stations a particular section of track is under
+maintenance, forcing trains going both directions to share a track. These basic issues result in a couple of clear
+concepts that we'll incorporate in our model.
+
+The first is a train. A train encapsulates all of the information relevant to our application: which line it is
+(Orange, Blue, or Silver), it's direction, and it's location within the system. Measuring the throughput of trains
+in the system will be our primary statistic when running the simulation.
+
+The other important entity to be concerned with is the track itself. From a queueing standpoint, it is a limited
+resource that only a single train can use at a single point. If a piece of track is occupied by a train, all other
+trains must wait before they may proceed through the track (in an assumed FIFO order). It is for this reason that
+most train systems do prefer to have a second track to prevent blocking for trains going in opposite directions.
+So, in our case we can treat a section of track as a queue that can only process a single train at a time. Between
+stations, there are generally two sections of track. However, track maintenance would cause a track to close, forcing
+opposing tracks to share.
+
+Because of this, we can represent the entire metro as queuing network with sections between stations as the queue. In
+our case, we will only be modeling the section of track between Rosslyn and Stadium-Armory where the Orange, Blue, and
+Silver share a track (See \ref{fig:metromap}). Since there are no track crossovers, we can consider these three lines
+to be isolated from other lines.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=5in]{../images/metro_map.pdf}
+\caption{\small \sl WMATA Metrorail Map \label{fig:metromap} \cite{fioroni}}
+\end{center}
+\end{figure}
 
 \section{Data Collection}
+
+\subsection{Query Design}
+Viewing this as a queueing problem, there are some ``missing pieces'' that our model did not incorporate, namely the
+arrival rate of trains and the service time for a single train. Instead of attempting to calculate these based on 
+numerical models of motion, we can instead use the WMATA API \cite{wmataapi} to calculate this.
+
+The API provides data by station with an estimate at each station of the wait time until the next train. For instance,
+the API for a station on the blue line might list a train to Franconia-Springfield (westbound) in 1 minute. If there
+also happens to be an Orange that passes through that station, you may also see an estimate for the next Orange (if it
+close enough).
+
+Given this information, we can calculate how often a train runs simply by marking the timestamp of the API call and
+offset by the estimated time until arrival provided by the API. For instance, if we query a station at 12:00 PM, and
+the API estimates a train arriving in 3 minutes. We can roughly estimate that the train will arrive at 12:03. This
+gives us our arrival rate Furthermore, we can query by train line as well to verify that trains run according to the
+schedule that WMATA provides on its website \cite{metroschedule}.
+
+To calculate service times for a particular section of track, we need an estimate of the time it takes for a train to
+travel between two stations. This is clearly station dependent, since some stations are closer than others. However,
+all but one of our stations are within the District of Columbia, so we expect the travel times to be small.
+
+We can still calculate an estimated travel time through the API as well. Given two neighboring stations $A$ and $B$, if
+we query $A$ for an upcoming trains of color $c$ in the direction of $B$, we will get some estimate for the arrival. We
+can query $B$ for the same color train in the same direction to obtain a second estimate. Subtracting these two numbers
+provides an estimate of the travel time between stations. This is making an assumption that there is no train of the
+same color in the middle that may lower the estimate. However, if we restrict our queries on $A$ to trains that are only
+a minute away, we are unlikely to have that problem since the stations are so close.
+
+\subsection{Query Results}
 
@@
 
-\section{Remaining Work}
+\section{Current and Remaining Work}
 

### 2015-03-31 project2/paper/main.tex
Fleshing out the data section even more

@@
 \subsection{Query Results}
+A script was built to call the WMATA API for train estimates, and was setup as a cron job on server. The script
+essentially downloaded the estimates of train arrivals for every station into a single timestamped JSON file (see the
+JSON format \cite{json}. Cron \cite{crontab} was configured to run the script every minute for an entire week,
+resulting in a directory full of timestamped JSON files. These were then imported into a Mongo \cite{mongodb} database 
+to allow easier queries.
+
+Train interarrivals were measured by line and direction. \ref{fig:eastboundblueinterarrivals} demonstrates a case where
+the data appears to be normal. However, other cases such as \ref{fig:westboundsilverinterarrivals} show that this is
+not necessarily true. Graphs for the remaining directions and line colors were omitted for brevity, but are included in
+the submission. Due to time constraints, we'll use an empirical distribution rather than trying to fit anything.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4in]{../images/train_interarrivals/eastbound_blue_interarrivals.png}
+\caption{\small \sl Eastbound Blue Line Interarrivals at Rosslyn \label{fig:eastboundblueinterarrivals}}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4in]{../images/train_interarrivals/westbound_silver_interarrivals.png}
+\caption{\small \sl Westbound Silver Line Interarrivals at Stadium Armory \label{fig:westboundsilverinterarrivals}}
+\end{center}
+\end{figure}
 

### 2015-03-31 project2/paper/main.tex
Getting the mean for the blue line
Adding statistics for Silver

@@
 \includegraphics[width=4in]{../images/train_interarrivals/eastbound_blue_interarrivals.png}
-\caption{\small \sl Eastbound Blue Line Interarrivals at Rosslyn \label{fig:eastboundblueinterarrivals}}
+\caption{\small \sl Eastbound Blue Line Interarrivals at Rosslyn. Ranges from 2.71 minutes to 16.78 with mean 10.91 minutes \label{fig:eastboundblueinterarrivals}}
 \end{center}
@@
 \includegraphics[width=4in]{../images/train_interarrivals/westbound_silver_interarrivals.png}
-\caption{\small \sl Westbound Silver Line Interarrivals at Stadium Armory \label{fig:westboundsilverinterarrivals}}
+\caption{\small \sl Westbound Silver Line Interarrivals at Stadium-Armory. Ranges from 2 minutes to 24 with mean 9.25 minutes \label{fig:westboundsilverinterarrivals}}
 \end{center}

### 2015-03-31 project2/paper/main.tex
Filling out the caveats section

@@
 
+Time between stations results were fairly predictable, and most stations averaged around a 2 minute travel time between
+stations. However, between Foggy Bottom and Rosslyn the times instead average around a minute higher. This is simply
+due to the distance between the stations, which is considerably greater between those stations.
+
 \section{Caveats and Limitations}
+With any sort of simulation there are simplications and assumptions that are made, and this project is of course no
+different. In some cases the reduction in detail does not harm accuracy but simply makes the model designer's and
+programmer's life easier. However, sometimes this is not the case and the simplifications greatly reduce the accuracy
+of the model. Thus, it is the modeler's job to judiciously choose which details to omit and leave in when building the
+model.
+
+One of the simplifications that we've made is that we've excluded all other stations in the metro system except those
+between Rosslyn and Stadium Armory. Other lines, the Red, Green, and Yellow lines, do not share track with any train,
+nor do they cross at any point. In this respect, they are isolated (regardless of the fact that they may share
+stations). However, ignoring Blue, Orange, and Silver stations beyond Stadium-Armory and Rosslyn could potentially
+reduce model accuracy here because train events at those stations can very easily influence the stations within D.C.
+We have chosen instead to model those exterior stations as a probability distribution which is based off of empirical
+data, but we still run a risk.
+
+That being said, choosing an empirical distribution like we have has the potential to be a wise or poor choice. If the
+"true distribution" of train interarrivals and between station times abides by what we have observed, then our choice
+of empirical distribution was smart, and we've saved ourselves some effort. However we don't know the true
+distribution, and it could be that our observations do not match that distribution at all.
 

### 2015-03-31 project2/paper/main.tex
Adding a little more to the caveat section

@@
 
+For individual trains, we'll also ignore some of the micro-effects of acceleration and deceleration and assume our
+between station times measurements captures that. For visualization purposes, we can simply draw trains as moving at a
+constant velocity through the system, although this is clearly not the case.
+
 \section{Technology Used}
+The simulation was built using CoffeeScript \cite{coffeescript}, using Backbone \cite{backbone}, Underscore
+\cite{underscore} and jQuery \cite{jQuery} to organize code some. Code was bundled via Browserify \cite{browserify},
+and minified with UglifyJS \cite{uglify}. Gulp \cite{gulp} was used as a build tool and package management was done
+via NPM \cite{npm}.
 

### 2015-03-31 project2/sim/gulpfile.js
No uglify right now

@@ -6,7 +6,7 @@ var uglify = require('gulp-uglify');
 gulp.task('bundle', function() {
     return gulp.src('coffee/main.coffee')
         .pipe(coffeeify())
-        .pipe(uglify({ "source-map" : "main.js.map" }))
+        //.pipe(uglify({ "source-map" : "main.js.map" }))
         .pipe(gulp.dest('.'));
 });
 

### 2015-03-31 project2/paper/main.tex
Adding the final section

@@
 and minified with UglifyJS \cite{uglify}. Gulp \cite{gulp} was used as a build tool and package management was done
-via NPM \cite{npm}.
+via NPM \cite{npm}. Geo-mapping was done via Leaflet \cite{leaflet}.
 
 \section{Current and Remaining Work}
+Currently, the app is simply mapping stations and labeling them. The entirety of data analysis is complete, and can be
+viewed within the \texttt{queries} directory, along with the \texttt{utils}. A few data files are strewn about the
+codebase as well. The WMATA API query code is stored within the \texttt{wmata} folder.
+
+I have also copied some code over from the first project since it will be applicable, such as a discrete event queue.
+However, the \texttt{Track} class remains to be written and is integral to the application. Furthermore, the actual
+train scheduling algorithm itself needs to be written.
 

### 2015-03-31 project2/paper/main.tex
Adding a reference to the application
Finalizing the current draft

@@
 subways. Commercial software such as OpenTrack \cite{opentrack} have been developed, and even video games such as Train
-Simulator \cite{trainsimulator} and Open Rails \cite{openrails}. Algorithms for handling single, double and tracks exist, such as the basic intuitive
+Simulator \cite{trainsimulator} and Open Rails \cite{openrails}. Algorithms for handling single, double and triple tracks exist,
+such as a basic
 algorithm for sharing tracks between trains presented by Fiorini and Botter \cite{fioroni}. Dessouky and Leachman
@@
 Silver share a track (See \ref{fig:metromap}). Since there are no track crossovers, we can consider these three lines
-to be isolated from other lines.
+to be isolated from other lines. However, we do need to keep in mind that for two tracks, each track is devoted to a
+specific direction. Only in the case of a single track is when a track will handle both directions.
+
+Because this will be a discrete-event simulation, we will need to be mindful of the events that travel through our
+application. A first example that comes to mind is the \texttt{train:arrived} event, which will be received by the
+Rosslyn and Stadium-Armory stations when a train arrives. This event will need to contain which line, along with the
+direction its heading (although that is implied by the station in our case).
+
+Track sections will need to inform neighboring stations when they are free, or if they are occupied. This of course
+will influence when and where trains are sent in order to reach their destination. In this respect, track sections are
+the queues in this network.
 
@@
 also happens to be an Orange that passes through that station, you may also see an estimate for the next Orange (if it
-close enough).
+is close enough).
 
@@
 offset by the estimated time until arrival provided by the API. For instance, if we query a station at 12:00 PM, and
-the API estimates a train arriving in 3 minutes. We can roughly estimate that the train will arrive at 12:03. This
-gives us our arrival rate Furthermore, we can query by train line as well to verify that trains run according to the
+the API estimates a train arriving in 3 minutes, we can roughly estimate that the train will arrive at 12:03. This
+gives us our arrival rate. Furthermore, we can query by train line as well to verify that trains run according to the
 schedule that WMATA provides on its website \cite{metroschedule}.
@@
 
-We can still calculate an estimated travel time through the API as well. Given two neighboring stations $A$ and $B$, if
+We can still calculate an estimated travel time between two stations through the API as well. Given two neighboring stations $A$ and $B$, if
 we query $A$ for an upcoming trains of color $c$ in the direction of $B$, we will get some estimate for the arrival. We
-can query $B$ for the same color train in the same direction to obtain a second estimate. Subtracting these two numbers
-provides an estimate of the travel time between stations. This is making an assumption that there is no train of the
-same color in the middle that may lower the estimate. However, if we restrict our queries on $A$ to trains that are only
-a minute away, we are unlikely to have that problem since the stations are so close.
+can query $B$ for the same color train in the same direction to obtain a estimate of the arrival time at $B$.
+Subtracting these two numbers provides an estimate of the travel time between stations. This makes an assumption that
+there is no train of the same color in the middle that may skew the estimate. However, we can minimize the chance of
+this happening if we restrict our queries on $A$ to trains that are only a minute away. Given that the stations are so
+close, we minimize the likelihood that there is a ``middle train'' that causes inaccurate estimates.
 
@@
 not necessarily true. Graphs for the remaining directions and line colors were omitted for brevity, but are included in
-the submission. Due to time constraints, we'll use an empirical distribution rather than trying to fit anything.
+the submission. Due to time constraints, we'll use an empirical distribution rather than trying to fit any particular
+distribution.
 
@@
 \section{Caveats and Limitations}
-With any sort of simulation there are simplications and assumptions that are made, and this project is of course no
+With any sort of simulation of a real-world system, there are simplications and assumptions that are made, and this project is of course no
 different. In some cases the reduction in detail does not harm accuracy but simply makes the model designer's and
@@
 nor do they cross at any point. In this respect, they are isolated (regardless of the fact that they may share
-stations). However, ignoring Blue, Orange, and Silver stations beyond Stadium-Armory and Rosslyn could potentially
+stations). So, in this case, the simplification hinders accuracy in no way whatsoever. However, ignoring Blue, Orange,
+and Silver stations beyond Stadium-Armory and Rosslyn could potentially
 reduce model accuracy here because train events at those stations can very easily influence the stations within D.C.
 We have chosen instead to model those exterior stations as a probability distribution which is based off of empirical
-data, but we still run a risk.
+data, but we still run a risk that we may have not captured an important interaction.
 
@@
 \section{Current and Remaining Work}
-Currently, the app is simply mapping stations and labeling them. The entirety of data analysis is complete, and can be
+Currently, the app is simply mapping stations and labeling them based on information provided by the API. The entirety
+of data analysis is complete, and can be
 viewed within the \texttt{queries} directory, along with the \texttt{utils}. A few data files are strewn about the
@@
 I have also copied some code over from the first project since it will be applicable, such as a discrete event queue.
-However, the \texttt{Track} class remains to be written and is integral to the application. Furthermore, the actual
-train scheduling algorithm itself needs to be written.
+However, the \texttt{Track} class remains to be written and is integral to the application, along with the \texttt{Train}
+class. Furthermore, the actual
+train scheduling algorithm itself needs to be written. However, the current application is deployed here
+\cite{aprilandchip}.
 

### 2015-04-19 project2/sim/gulpfile.js
Minifying again now

@@ -7,7 +7,7 @@ var uglify = require('gulp-uglify');
 gulp.task('bundle', function() {
     return gulp.src('coffee/main.coffee')
         .pipe(coffeeify())
-        //.pipe(uglify({ "source-map" : "main.js.map" }))
+        .pipe(uglify({ "source-map" : "main.js.map" }))
         .pipe(gulp.dest('.'));
 });
 

### 2015-04-20 project2/paper/gulpfile.js
Adding a watch task for the gulp file

@@ -26,3 +26,7 @@ gulp.task('default', ['clean'], function(cb) {
         asyncCommand(latex)
     ], cb);
 });
+
+gulp.task('watch', ['default'], function() {
+    return gulp.watch(['main.bib', 'main.tex'], ['default']);
+});

### 2015-04-22 project2/paper/main.tex
Fleshing out a few things

@@
 
-The first is a train. A train encapsulates all of the information relevant to our application: which line it is
+The first is a train. A \texttt{Train} encapsulates all of the information relevant to our application: which line it is
 (Orange, Blue, or Silver), it's direction, and it's location within the system. Measuring the throughput of trains
@@
 
-The other important entity to be concerned with is the track itself. From a queueing standpoint, it is a limited
+The other important entity to be concerned with is the track itself. In the simulation, this is modeled using a
+\texttt{TrackSegment} class. A \texttt{TrackSegment} is simply a queue with the ability to be process \texttt{Train}s
+one at time.
+
+From a queueing standpoint, a \texttt{TrackSegment} is a limited
 resource that only a single train can use at a single point. If a piece of track is occupied by a train, all other
@@
 
+I've also chosen to conceptually model pairs of tracks between stations as a \texttt{StationConnection} class. Perhaps
+unsurprisingly, this is the largest and most important class in the simulation, since it manage two
+\texttt{TrackSegment}s, along with the disabling/enabling of a track segment. When a track is disabled, the
+\texttt{StationConnection} must begin routing trains through a single track, rather than reserving tracks for
+only particular directions. This logic is important to the accuracy and realism of the simulation, since it's one of
+fundamental principles behind a train track.
+
 Because this will be a discrete-event simulation, we will need to be mindful of the events that travel through our
-application. A first example that comes to mind is the \texttt{train:arrived} event, which will be received by the
-Rosslyn and Stadium-Armory stations when a train arrives. This event will need to contain which line, along with the
-direction its heading (although that is implied by the station in our case).
+application. A first example that comes to mind is the \texttt{train:arrive} event, which will be received by the
+Rosslyn and Stadium-Armory stations when a train first arrives in the system. This event will need to contain the train
+instance, along with the direction its heading (although that is implied by the station in our case). This way, we can
+track the train as it moves through the system to calculate metrics about the system's performance.
 
-Track sections will need to inform neighboring stations when they are free, or if they are occupied. This of course
-will influence when and where trains are sent in order to reach their destination. In this respect, track sections are
-the queues in this network.
+When a train first arrives at a station, it is queued up to await a free track. If the track designated for that
+direction is already available, the train simply moves on to the track and occupies it until it reaches the next
+station where it exits and then queues up at the next station.
+
+When a train is dequeued and allowed to enter a track, a \texttt{train:enter} event is thrown. When this event occurs,
+the station connection marks the track as occupied so that later arriving trains do not attempt to enter the track. The
+connection also schedules a new event: \texttt{train:exit}
+
+\texttt{train:exit} is triggered when a train exits a connection, which informs a connection to free up the track and
+push the next train through if there is one. As mentioned before, this event is actually scheduled when a train enters
+a track between stations. Each connection has knowledge of the time between it's associated stations, and uses that
+time to schedule the \texttt{train:exit} in relation to \texttt{train:enter} for a particular train. Lastly, when a
+train exits a particular connection, a \texttt{train:arrive} event is scheduled immediately for the next connection in
+the metro system.
 
@@
 and minified with UglifyJS \cite{uglify}. Gulp \cite{gulp} was used as a build tool and package management was done
-via NPM \cite{npm}. Geo-mapping was done via Leaflet \cite{leaflet}.
+via NPM \cite{npm}. Geo-mapping was done via Leaflet \cite{leaflet}, which provides an API for enhancing the map with
+markers, text, shapes, etc.
+
+In an effort to ensure code is working properly, the simulation was built with a testing suite. Mocha \cite{mocha} was
+used as the testing framework, and Chai \cite{chai} was used for BDD-style assertions. Sinon \cite{sinon} was used to
+mock function calls during tests.
 

### 2015-04-24 project2/paper/main.tex
Adding some better headings
Fixing the date
Stubbing out the last section

@@
 \author{Chip Bell}
-\date{March 31, 2015}
+\date{April 24, 2015}
 \maketitle
@@
 
-\section{Current and Remaining Work}
-Currently, the app is simply mapping stations and labeling them based on information provided by the API. The entirety
-of data analysis is complete, and can be
-viewed within the \texttt{queries} directory, along with the \texttt{utils}. A few data files are strewn about the
-codebase as well. The WMATA API query code is stored within the \texttt{wmata} folder.
+\section{Current Build and Instructions}
+The functioning simulation is currently hosted at \cite{aprilandchip}. On this page, you can see trains moving through
+the metro system, with real-time statistics as trains reach the other side. Also, tracks can be disabled via the
+checkboxes at the bottom, which will influence trains in real-time. The map itself is interactive, and provides a few
+glimpses into the system. Clicking on a station gives you the name, and clicking on a connection tells you current stats
+at that section of track.
 
-I have also copied some code over from the first project since it will be applicable, such as a discrete event queue.
-However, the \texttt{Track} class remains to be written and is integral to the application, along with the \texttt{Train}
-class. Furthermore, the actual
-train scheduling algorithm itself needs to be written. However, the current application is deployed here
-\cite{aprilandchip}.
+Also within the project directory that has been uploaded with the submission are some of the data analysis tools I've
+built for performing queries, and cleaning data. These can be found in the \texttt{queries} and \texttt{utils}
+directory of the project. Among the scripts found include bash scripts for file templating (for building queries
+dynamically), templated js scripts for querying a mongo database, along with a few Python scripts for performing data
+cleanup (like boxcar filtering for smoothing, and simple binning for histograms).
+
+\section{Results and Conclusions}
+One of the goals for this project was to better understand the metro system as a whole. In particular, I wanted to
+discover how accurate WMATA was at predicting trains, and also how robust their scheduling was to unpredicted track
+unavailability. WMATA provides their own estimates of how often trains occur \cite{metroschedule}, along with between
+station times from any station to another \cite{rosslyntime}. This gives us plenty of room for comparison.
+
+% Talking points
+% - Average between station time
+% - Rosslyn to Stadium Armory time is consistent
+% - Robustness to changes in track availability
+
+Unsurprisingly, WMATA plans its trains effectively to be robust to a fair amount track outages, while still providing
+the timings estimated on their website.
 

### 2015-04-24 project2/paper/main.tex
Adding conclusion section

@@
 
-% Talking points
-% - Average between station time
-% - Rosslyn to Stadium Armory time is consistent
-% - Robustness to changes in track availability
+Empirically measuring between station times is a very granular way to measure the accuracy of WMATA's predictions. For
+instance, stations within DC are estimated between 1 and 2 minutes, resulting in a travel time from Rosslyn to
+Stadium-Armory taking 21 minutes. Based on the between station data we collected, we estimate empirically this actaully
+ends up being around 24 minutes and 8 seconds. This discrepancy can be explained by a few reasons, the first being that
+WMATA is wrong. This is of course, a potential case, but I'm more sure that this really an indication of issues in the
+dataset itself. Since rail predictions are only precise to the minute, stations that are only a minute apart have the
+potential to cause aliasing when calculating between station times. This may be the reason some of the interior station
+times are 2 minutes instead of 1. Moreover, because this was a global measurement over a period of time, I may have 
+captured some delaying event that caused between station times to diminish to a great enough value for a long enough
+time to ``pull'' the mean up.
+
+Regardless of these discrepancies, the simulation appears to agree with the estimates. Trains moving westward average
+around 22.15 minutes to reach Rosslyn from Stadium-Armory. Eastward trains perform slightly better, hovering closer to
+21 minutes. Both distributions, without any track disturbances, are very tight with little variance. The difference in
+means is accounted for by the empirical measurement of between station times. The total westward time from Rosslyn to
+Stadium-Armory is roughly 30 seconds shorter than the eastward counterpart. The remaining time may be accounted for by
+the Rosslyn-Foggy Bottom connection, the most western connection in the system. The between station time here is 
+consistently 3 minutes, rather than 2. Because of this, arriving trains may be forced to wait more often. However,
+because our scope is narrowed for this project, we may be missing some other behavior further westward that may explain
+this behavior.
 
 Unsurprisingly, WMATA plans its trains effectively to be robust to a fair amount track outages, while still providing
-the timings estimated on their website.
+the timings estimated on their website. Even with three tracks disabled, the mean travel time of eastern and western
+trains still hovers around 23 minutes and 30 seconds. However, worth noting is that the \emph{variabilty} increases
+drastically. Even though the mean moves little, there still are some trains that end up taking over 40 minutes to
+reach the other side given these conditions. In a way, this is simply ``bad luck'' on that particular train's part,
+simply because it's timing fell so that it arrived right as another train had occupied the track (with potentially
+other trains waiting already). However as more tracks become disabled, the bad luck slowly becomes more likely for any
+train to experience.
 

### 2015-04-25 project2/paper/main.tex
Final cleanup!

@@
 our case, we will only be modeling the section of track between Rosslyn and Stadium-Armory where the Orange, Blue, and
-Silver share a track (See \ref{fig:metromap}). Since there are no track crossovers, we can consider these three lines
-to be isolated from other lines. However, we do need to keep in mind that for two tracks, each track is devoted to a
-specific direction. Only in the case of a single track is when a track will handle both directions.
+Silver share a track (See \ref{fig:metromap}) which results in a simpler network with the nodes falling in a single
+line. Since there are no track crossovers within the DC metro, we can consider these three lines
+to be isolated from other lines, so we lose no accuracy by excluding the other lines from the simulation. However, we
+do need to keep in mind that for two tracks, each track is devoted to a specific direction. Only in the case of a single
+track is when a track will handle both directions.
 
@@
 Rosslyn and Stadium-Armory stations when a train first arrives in the system. This event will need to contain the train
-instance, along with the direction its heading (although that is implied by the station in our case). This way, we can
+instance, along with the direction its heading. This way, we can
 track the train as it moves through the system to calculate metrics about the system's performance.
@@
 \texttt{train:exit} is triggered when a train exits a connection, which informs a connection to free up the track and
-push the next train through if there is one. As mentioned before, this event is actually scheduled when a train enters
+push the next train through, if there is one. As mentioned before, this event is actually scheduled when a train enters
 a track between stations. Each connection has knowledge of the time between it's associated stations, and uses that
@@
 arrival rate of trains and the service time for a single train. Instead of attempting to calculate these based on 
-numerical models of motion, we can instead use the WMATA API \cite{wmataapi} to calculate this.
+numerical models of motion, we can instead use the WMATA API \cite{wmataapi} to claculate an empirical value.
 
@@
 offset by the estimated time until arrival provided by the API. For instance, if we query a station at 12:00 PM, and
-the API estimates a train arriving in 3 minutes, we can roughly estimate that the train will arrive at 12:03. This
-gives us our arrival rate. Furthermore, we can query by train line as well to verify that trains run according to the
-schedule that WMATA provides on its website \cite{metroschedule}.
+the API estimates a train arriving in 3 minutes, we can roughly estimate that the train will arrive at 12:03.
+Performing this operation over time gives us a list of train arrival time by station from which we can calculate
+interarrival time via a running diff. Furthermore, we can query by train line as well to verify that trains run
+according to the schedule that WMATA provides on its website \cite{metroschedule}.
 
@@
 \subsection{Query Results}
-A script was built to call the WMATA API for train estimates, and was setup as a cron job on server. The script
+A script was built to call the WMATA API for train estimates, and was setup as a cron job on a server and scheduled to
+run every minute for a week. The script
 essentially downloaded the estimates of train arrivals for every station into a single timestamped JSON file (see the
 JSON format \cite{json}. Cron \cite{crontab} was configured to run the script every minute for an entire week,
-resulting in a directory full of timestamped JSON files. These were then imported into a Mongo \cite{mongodb} database 
+resulting in a directory full of timestamped JSON files. These were then imported into a MongoDB \cite{mongodb} database 
 to allow easier queries.
 
-Train interarrivals were measured by line and direction. \ref{fig:eastboundblueinterarrivals} demonstrates a case where
-the data appears to be normal. However, other cases such as \ref{fig:westboundsilverinterarrivals} show that this is
-not necessarily true. Graphs for the remaining directions and line colors were omitted for brevity, but are included in
+Train interarrivals were measured by line and direction. (\ref{fig:eastboundblueinterarrivals}) demonstrates that the
+data oftentimes appears to follow some ``standard'' distribution. However, other cases such as
+(\ref{fig:westboundsilverinterarrivals}) show that this is not necessarily true. Graphs for the remaining directions
+and line colors were omitted for brevity, but are included in
 the submission. Due to time constraints, we'll use an empirical distribution rather than trying to fit any particular
@@
 between Rosslyn and Stadium Armory. Other lines, the Red, Green, and Yellow lines, do not share track with any train,
-nor do they cross at any point. In this respect, they are isolated (regardless of the fact that they may share
-stations). So, in this case, the simplification hinders accuracy in no way whatsoever. However, ignoring Blue, Orange,
-and Silver stations beyond Stadium-Armory and Rosslyn could potentially
+nor do they cross at any point. In this respect, they are isolated. Event though other lines may share stations, the
+trains run at a different elevation and therefore do no share a track. So, in this case, the simplification hinders
+accuracy in no way whatsoever. However, ignoring Blue, Orange, and Silver stations beyond Stadium-Armory and Rosslyn
+could potentially
 reduce model accuracy here because train events at those stations can very easily influence the stations within D.C.
@@
 \section{Technology Used}
-The simulation was built using CoffeeScript \cite{coffeescript}, using Backbone \cite{backbone}, Underscore
+The simulation was built using CoffeeScript \cite{coffeescript}, Backbone \cite{backbone}, Underscore
 \cite{underscore} and jQuery \cite{jQuery} to organize code some. Code was bundled via Browserify \cite{browserify},
-and minified with UglifyJS \cite{uglify}. Gulp \cite{gulp} was used as a build tool and package management was done
-via NPM \cite{npm}. Geo-mapping was done via Leaflet \cite{leaflet}, which provides an API for enhancing the map with
+and minified with UglifyJS \cite{uglify}. Gulp \cite{gulp} was used as a build tool and NPM \cite{npm} was used as a
+package manager. Geo-mapping was provided by Leaflet \cite{leaflet}, which has an API for enhancing the map with
 markers, text, shapes, etc.
 
-In an effort to ensure code is working properly, the simulation was built with a testing suite. Mocha \cite{mocha} was
-used as the testing framework, and Chai \cite{chai} was used for BDD-style assertions. Sinon \cite{sinon} was used to
-mock function calls during tests.
+In an effort to ensure that the code worked properly, the simulation was built with a testing suite. Mocha \cite{mocha}
+was used as the testing framework, and Chai \cite{chai} was used for BDD-style assertions. Sinon \cite{sinon} was used
+to mock function calls during tests.
 
@@
 Regardless of these discrepancies, the simulation appears to agree with the estimates. Trains moving westward average
-around 22.15 minutes to reach Rosslyn from Stadium-Armory. Eastward trains perform slightly better, hovering closer to
-21 minutes. Both distributions, without any track disturbances, are very tight with little variance. The difference in
-means is accounted for by the empirical measurement of between station times. The total westward time from Rosslyn to
-Stadium-Armory is roughly 30 seconds shorter than the eastward counterpart. The remaining time may be accounted for by
-the Rosslyn-Foggy Bottom connection, the most western connection in the system. The between station time here is 
-consistently 3 minutes, rather than 2. Because of this, arriving trains may be forced to wait more often. However,
-because our scope is narrowed for this project, we may be missing some other behavior further westward that may explain
-this behavior.
+around 22.15 minutes to reach Rosslyn from Stadium-Armory within the simulation. Eastward trains perform slightly
+better, hovering closer to 21 minutes. Both distributions, without any track disturbances, are very tight with little
+variance. The difference in  means is accounted for by the empirical measurement of between station times. The total
+westward time from Rosslyn to  Stadium-Armory is roughly 30 seconds shorter than the eastward counterpart. This could
+be a result of many factors, such as geography, or even interactions beyond Rosslyn and Stadium-Armory ``captured''
+in our empirical distribution. The remaining time may be accounted for by the Rosslyn-Foggy Bottom connection, the
+most western connection in the system. The between station time here is  consistently 3 minutes, rather than 2.
+Because of this, arriving trains may be forced to wait more often. WMATA estimates the travel time being the same
+regardless of direction, and I still cannot explain the discrepancy.
 
@@
 other trains waiting already). However as more tracks become disabled, the bad luck slowly becomes more likely for any
-train to experience.
+train to experience. Eventually, the mean does increase to around 32 minutes, with a very wide distribution with many
+trains exceeding 45 minute travel times.
+
+In conclusion, we see that WMATA does indeed provide fairly accurate estimates of train travel times, along with the
+interarrival times between stations. Furthermore, WMATA's train scheduling is fairly robust to track outages, without
+drastically impacting the throughput of trains. There are some unanswered questions, however. Why does the direction
+affect the travel time between stations? Why do the simulated travel times agree with WMATA's estimate, but the average
+interarrivals measured from the API do not? Would increasing the scope of the model improve the accuracy and remove the
+discrepancies altogether? These questions can perhaps be answered with deeper investigation and more data.
 
